### 2017-11-24 docs/report/draft.tex
Started the report.

\documentclass[10pt,a4paper]{article}

\usepackage[utf8]{inputenc}

% Information for header.
\title{Optimizing the Cobb Doughlas Habitability Production function using
Particle Swarm Optimization}
\author{Abhijit J.\ Theophilus}
\date{\today}

% Useful for initial drafts.
\newenvironment{pointers}{%
  \noindent Should include,
  \begin{itemize}
    \setlength{\itemsep}{-1pt}}{%
\end{itemize}}

\begin{document}
\maketitle

\begin{abstract}
  Write a super cool abstract here.
\end{abstract}

\section{Introduction}
What am I trying to do?\\

\begin{pointers}
\item What is CDHPF?\ Why use a metric?
\item Previous attempts.
\item What is PSO?\
\item Why PSO?\
\item Overview of paper.
\end{pointers}

\section{Cobb Douglas Habitability Score}
\begin{pointers}
\item A little background (expanding from Introduction).
\item The equations.
\item Method for calculating the score.
\end{pointers}

\section{Particle Swarm Optimization}
\begin{pointers}
\item Equations. Modification for max.
\item Constrained optimization modifications.
\item Any other changes made.
\item Algorithm: Putting everything together.
\end{pointers}

\section{Experiment}
\begin{pointers}
\item Discussing the data set.
\item Parameters for the constrained optimization.
\item Implementation method.
\item Ensuring convergence.
\end{pointers}

\section{Results}
\begin{pointers}
\item The values and proximity to earth's habitability score.
\item Values that do not converge. Or were hard to converge.
\item Speed of convergence graphs.
\item Graph1: Iterations to convergence vs. Number of particles.
\item Graph2: Distribution of number of iterations to convergence.
\item Graph3: Iterations to convergence vs. Constraint parameters.
\end{pointers}

\section{Conclusions}
\begin{pointers}
\item Why is the speed so important?
\item Parallelizable.
\end{pointers}

\end{document}

### 2017-11-26 cdhs_by_pso.py
Changed the way the code is run.

@@ -75,54 +75,67 @@ tot = len(exoplanets)
 
 results = {'CRS': [], 'DRS': []}
 
-# Estimating CDHS at CRS.
-print('')
-for constraint in ['CRS', 'DRS']:
-    results = []
-    results.append(['Name', 'A', 'B', 'CDHSi', 'G', 'D', 'CDHSs',
-                    'CDHS', 'Cls'])
+
+def evaluate_cdhs_values():
+    global sw_kwargs
+    print('')
+    for constraint in ['CRS', 'DRS']:
+        results = []
+        results.append(['Name', 'A', 'B', 'CDHSi', 'G', 'D', 'CDHSs',
+                        'CDHS', 'Cls'])
+
+        print('=' * 82, end='\n\n')
+        print('#', constraint, end='\n\n')
+        print(msg.format(*results[-1]))
+        print('-' * 82)
+
+        curr = 1
+        for _, row in exoplanets.iterrows():
+            # CDHSi
+            cdhpf_i = construct_cdhs_fn(row['Radius'], row['Density'],
+                                        constraint)
+            swarm_i = converge_by_pso(fn=cdhpf_i, **sw_kwargs)
+            if not swarm_i:
+                print(err.format(row['Name'],
+                                 '** CDHSi convergence failed. **'))
+                curr += 1
+                continue
+
+            A, B = np.round(swarm_i.best_particle.best, 2)
+            cdhs_i = round(swarm_i.global_best, 4)
+
+            # CDHSs
+            row['STemp'] = row['STemp'] / 288           # Normalizing to EU.
+            cdhpf_s = construct_cdhs_fn(row['Escape'], row['STemp'],
+                                        constraint)
+            swarm_s = converge_by_pso(fn=cdhpf_s, **sw_kwargs)
+            if not swarm_s:
+                print(err.format(row['Name'],
+                                 '** CDHSs convergence failed. **'))
+                curr += 1
+                continue
+
+            G, D = np.round(swarm_s.best_particle.best, 2)
+            cdhs_s = round(swarm_s.global_best, 4)
+
+            cdhs = np.round(cdhs_i*.99 + cdhs_s*.01, 4)
+            results.append([row['Name'], A, B, cdhs_i, G, D, cdhs_s, cdhs,
+                            row['Habitable']])
+            print(msg.format(*results[-1]))
+
+            progress = (curr*72) // tot
[…]

### 2017-11-26 pso/pso.py

@@ -64,7 +64,7 @@ class Swarm:
                 self.best_particle = particle
                 self.global_best = particle.best_fitness
 
-    def converge(self, max_stable=10, max_iter=100, threshold=0.0001,
+    def converge(self, max_stable=50, max_iter=10000, threshold=0.0001,
                  verbose=True, proglen=20):
         """Find the maxima for the fitness function specified."""
         stable_count = 0

### 2017-11-26 cdhs_by_pso_multi.py
Completed the parallel iter program.

@@ -2,9 +2,21 @@
 
 import csv
 import functools
+import logging
+from multiprocessing import Lock, Process
 import numpy as np
+
 from exoplanets import exoplanets
 from pso import Swarm
+
+
+# Constants.
+HEADERS = ('Name', 'A', 'B', 'CDHSi', 'G', 'D', 'CDHSs', 'CDHS', 'Cls')
+
+
+# Setting up logging.
+logging.basicConfig(format='[%(asctime)s]  %(message)s', level=logging.INFO)
+
 
 # Parameters for the swarm.
 sw_kwargs = {
@@ -23,6 +35,17 @@ sw_kwargs = {
 exoplanets.dropna(how='any', inplace=True)
 
 
+# Utility functions.
+def safe_log(logfunc, lock, pname, message):
+    """Multiprocess safe logging."""
+    lock.acquire()
+    try:
+        logfunc(pname+'  '+message)
+    finally:
+        lock.release()
+
+
+# Functions to set up the penalty.
 def penalty(pos, type_):
     """The penalty function for the Cobb Douglas function."""
     e = .1
@@ -57,6 +80,7 @@ def construct_cdhs_fn(coeff1, coeff2, type_):
     return cdhpf
 
 
+# Functions to evaluate the CDHS values for each exoplanet.
 def converge_by_pso(restarts=3, **kwargs):
     """Wait for convergence by PSO."""
     for _ in range(restarts):
@@ -67,27 +91,19 @@ def converge_by_pso(restarts=3, **kwargs):
     return None
 
 
-# Message text.
-msg = '{:25}{:>5}{:>5}{:>10}{:>5}{:>5}{:>10}{:>10}{:>7.5}'
-err = '{:25}{:^57}'
-bar = '[{:72}]  ({:>3}%)'
-tot = len(exoplanets)
+def evaluate_cdhs_values(lock, pname, **sw_kwargs):
+    """Estimate the CDHS values of each planet for both, the CRS and
+    the DRS constraints.
+    """
+    global exoplanets
+    num_exopl = len(exoplanets)
 
-results = {'CRS': [], 'DRS': []}
-
-
-def evaluate_cdhs_values():
-    global sw_kwargs
-    print('')
+    exoplanets = exoplanets[:20]
     for constraint in ['CRS', 'DRS']:
+        safe_log(logging.info, lock, pname,
+                 'Commencing '+constraint+' estimation.')
         results = []
[…]

### 2017-11-26 pso_test.py

@@ -3,7 +3,7 @@
 import numpy as np
 from pso import Swarm
 
-swarm = Swarm(10, ndim=1, fitness_function=lambda x: -(x**2), friction=0.7,
+swarm = Swarm(10, ndim=1, fn=lambda x: -(x**2), friction=0.7,
               learn_rate1=.5, learn_rate2=.5, max_velocity=4.0)
 swarm.converge()
 print(np.round(swarm.best_particle.position, 4))

### 2017-11-26 cdhs_by_pso_multi.py
Now saves the number of iter too.

@@ -11,7 +11,8 @@ from pso import Swarm
 
 
 # Constants.
-HEADERS = ('Name', 'A', 'B', 'CDHSi', 'G', 'D', 'CDHSs', 'CDHS', 'Cls')
+HEADERS = ('Name', 'A', 'B', 'CDHSi', 'Conv_i',
+           'G', 'D', 'CDHSs', 'Conv_s', 'CDHS', 'Cls')
 
 
 # Setting up logging.
@@ -87,8 +88,8 @@ def converge_by_pso(restarts=3, **kwargs):
         swarm = Swarm(**kwargs)
         converged = swarm.converge(verbose=False)
         if converged:
-            return swarm
-    return None
+            return (converged, swarm)
+    return (0, None)
 
 
 def evaluate_cdhs_values(lock, pname, **sw_kwargs):
@@ -110,7 +111,7 @@ def evaluate_cdhs_values(lock, pname, **sw_kwargs):
             # CDHSi
             cdhpf_i = construct_cdhs_fn(row['Radius'], row['Density'],
                                         constraint)
-            swarm_i = converge_by_pso(fn=cdhpf_i, **sw_kwargs)
+            conv_i, swarm_i = converge_by_pso(fn=cdhpf_i, **sw_kwargs)
             if not swarm_i:
                 safe_log(logging.ERROR, lock, pname,
                          row['Name']+'did not converge for CDHSi.')
@@ -124,7 +125,7 @@ def evaluate_cdhs_values(lock, pname, **sw_kwargs):
             row['STemp'] = row['STemp'] / 288           # Normalizing to EU.
             cdhpf_s = construct_cdhs_fn(row['Escape'], row['STemp'],
                                         constraint)
-            swarm_s = converge_by_pso(fn=cdhpf_s, **sw_kwargs)
+            conv_s, swarm_s = converge_by_pso(fn=cdhpf_s, **sw_kwargs)
             if not swarm_s:
                 safe_log(logging.ERROR, lock, pname,
                          row['Name']+'did not converge for CDHSs.')
@@ -135,8 +136,8 @@ def evaluate_cdhs_values(lock, pname, **sw_kwargs):
             cdhs_s = round(swarm_s.global_best, 4)
 
             cdhs = np.round(cdhs_i*.99 + cdhs_s*.01, 4)
-            results.append([row['Name'], A, B, cdhs_i, G, D, cdhs_s, cdhs,
-                            row['Habitable']])
[…]

### 2017-11-26 pso/pso.py

@@ -76,7 +76,7 @@ class Swarm:
             if np.abs(old_best - self.global_best) < threshold:
                 stable_count += 1
                 if stable_count == max_stable:
-                    return True
+                    return it
             else:
                 stable_count = 0
             if verbose:

### 2017-11-26 cdhs_by_pso_multi.py
Ready for final execution.

@@ -99,7 +99,6 @@ def evaluate_cdhs_values(lock, pname, **sw_kwargs):
     global exoplanets
     num_exopl = len(exoplanets)
 
-    exoplanets = exoplanets[:20]
     for constraint in ['CRS', 'DRS']:
         safe_log(logging.info, lock, pname,
                  'Commencing '+constraint+' estimation.')

### 2017-11-27 cdhs_by_pso_multi.py
Fixed a bug in multi.

@@ -140,7 +140,7 @@ def evaluate_cdhs_values(lock, pname, **sw_kwargs):
 
             q, r = divmod(curr*100 // num_exopl, 10)
             if r == 0 and q != 0:
-                safe_log(logging.info, lock, pname, str(q)+'% complete.')
+                safe_log(logging.info, lock, pname, str(q*10)+'% complete.')
             curr += 1
 
         safe_log(logging.info, lock, pname,

### 2017-11-27 cdhs_by_pso_multi.py
Rearranged a few files.

@@ -145,8 +145,8 @@ def evaluate_cdhs_values(lock, pname, **sw_kwargs):
 
         safe_log(logging.info, lock, pname,
                  'Completed '+constraint+' estimation.')
-        fpath = 'res/pso_{0}_{1}.csv'.format(constraint.lower(),
-                                             sw_kwargs['npart'])
+        fpath = 'res/multiple/pso_{0}_{1}.csv'.format(constraint.lower(),
+                                                      sw_kwargs['npart'])
         with open(fpath, 'w') as resfile:
             csv.writer(resfile).writerows(results)
         safe_log(logging.info, lock, pname, 'Written results.')

### 2017-11-27 generate_plots.py
Finished generating plots.

@@ -1,0 +1,49 @@
+#!/usr/bin/python
+
+# import matplotlib as mpl
+from matplotlib import pyplot as plt
+import numpy as np
+import pandas as pd
+
+
+def getdists(const):
+    dists = []
+    for n in range(50, 150, 10):
+        fpath = 'pso_'+const+'_'+str(n)+'.csv'
+        cdhs = pd.read_csv('res/multiple/'+fpath)
+        dists.append((np.histogram(cdhs['Conv_i'], bins=range(48, 60))[0],
+                      np.histogram(cdhs['Conv_s'], bins=range(48, 60))[0]))
+    return dists
+
+
+def plotgraphs(plots, label):
+    fig, axes = plt.subplots(2, 5, sharex=True, sharey=True)
+
+    fig.text(0.5, 0.95, label, ha='center', size='large')
+    fig.text(0.5, 0.04, 'Iterations to Convergence', ha='center', size='small')
+    fig.text(0.07, 0.31, 'Count', va='center', rotation='vertical',
+             size='small')
+    fig.text(0.07, 0.73, 'Count', va='center', rotation='vertical',
+             size='small')
+    axes = axes.flatten()
+
+    npart = 50
+    for i, dists in enumerate(plots):
+        axes[i].plot(range(0, 11), dists[0], '-', linewidth=1.0)
+        axes[i].plot(range(0, 11), dists[1], '--', linewidth=1.0)
+        axes[i].set_title(str(npart)+' Particles', {'fontsize': 'medium'})
+        axes[i].tick_params(axis='both', labelsize='small')
+        npart += 10
+
+    axes[0].set_xticks(range(0, 11, 2))
+    fig.legend(axes[0].get_lines(), ('$CDHS_i$', '$CDHS_s$'))
+
+
+if __name__ == '__main__':
+    crs = getdists('crs')
+    plotgraphs(crs, 'Constant Returns to Scale')
+
+    drs = getdists('drs')
+    plotgraphs(drs, 'Decreasing Returns to Scale')
+
+    plt.show()

### 2017-11-27 docs/report/draft.tex
Added pointers to draft.
Finished introduction in report.

@@
-\documentclass[10pt,a4paper]{article}
+\documentclass{article}
 
 \usepackage[utf8]{inputenc}
+\usepackage{mathtools}
 
@@
 
+% Useful replacements.
+\newcommand{\pso}{Particle Swarm Optimization}
+
 \begin{document}
@@
 
-\begin{abstract}
-  Write a super cool abstract here.
-\end{abstract}
+\section{Introduction}
+% Note to self, improve the flow of the report.
+% Find citations.
+Under most paradigms, optimization of a continuous function requires the calculation of a gradient. This might not
+always be feasible for non-polynomial functions in high-dimensional search spaces. This is further complicated when the
+optimization must be performed under constraints. \pso\ is a metaheuristic for continuous function optimization that does
+away with the need for a gradient. It employs a large number of particles that traverse the search space converging
+toward a global best solution encountered by at least one of the particles.
 
-\section{Introduction}
-What am I trying to do?\\
+\pso\ is a distributed method that requires simple mathematical operators and few lines of code, making it a lucrative
+solution where computation resources are at a premium. Its implementation is highly parallelizable. It scales with the
+dimensionality of the search space. Constrained optimization can be incorporated within the standard algorithm through
+the use of a penalty function.
 
-\begin{pointers}
-\item What is CDHPF?\ Why use a metric?
-\item Previous attempts.
-\item What is PSO?\
-\item Why PSO?\
-\item Overview of paper.
-\end{pointers}
+% Pull out citations from the Applications paper.
+\pso\ has been successfully adapted to a wide range of design optimization problems, e.g., networks and VLSI design.
+\pso\ has also found applications in machine learning under clustering, feature detection and classification.  As a
+modelling paradigm, it has been used for constructing customer satisfaction models, MIDI music models, friction models,
+etc.
 
-\section{Cobb Douglas Habitability Score}
-\begin{pointers}
-\item A little background (expanding from Introduction).
-\item The equations.
-\item Method for calculating the score.
-\end{pointers}
+In this paper, \pso\ has been applied to estimate the Cobb Douglas Habitability Score of an exoplanet. The score is
+estimated by maximizing the Cobb-Douglas Habitability Production Function. The function considers four parameters,
+radius, mass, escape velocity and surface temperature, taken from the Exoplanets Catalog hosted by the Planetary
+Habitability Laboratory, UPR Arecibo. Since the Cobb-Douglas function is concave under certain constraints, the
+problem can be reduced to a convex optimization over the negative of the production function.
+
+\section{Cobb-Douglas Habitability Score}
+
+\subsection{Cobb-Douglas Habitability Production Function}
+The general form of the Cobb-Douglas production function is given by,
+\begin{equation}
+  Y = k\cdot{(x_1)}^\alpha\cdot{(x_2)}^\beta,
+\end{equation}
+where $k$ is a constant that may be set according to requirement, $x_1$ and $x_2$ are the input parameters to the
+function, $Y$ is the total output production value, and $\alpha$ and $\beta$ are called the elasticity coefficients.
+
+% \begin{pointers}
+% \item Describe the Cobb Douglas Habitability Score.
+% \item Describe CRS, DRS and IRS\@.
+% \item Splitting the CDHS score into two values.
+% \item State the equations.
+% \item Score is estimated by maximizing the CDHPF fn.
+% \end{pointers}
 
@@
 \item Constrained optimization modifications.
-\item Any other changes made.
-\item Algorithm: Putting everything together.
+\item Algorithm for PSO\@.
 \end{pointers}

### 2017-11-27 docs/report/draft.tex
Added content to report.

@@
 
-In this paper, \pso\ has been applied to estimate the Cobb Douglas Habitability Score of an exoplanet. The score is
-estimated by maximizing the Cobb-Douglas Habitability Production Function. The function considers four parameters,
-radius, mass, escape velocity and surface temperature, taken from the Exoplanets Catalog hosted by the Planetary
-Habitability Laboratory, UPR Arecibo. Since the Cobb-Douglas function is concave under certain constraints, the
-problem can be reduced to a convex optimization over the negative of the production function.
+In this paper, \pso\ has been applied to estimate the Cobb Douglas Habitability Score (CDHS) of an exoplanet. The score
+is estimated by maximizing the Cobb-Douglas Habitability Production Function (CD-HPF). The function considers four
+parameters, radius, mass, escape velocity and surface temperature, taken from the Exoplanets Catalog hosted by the
+Planetary Habitability Laboratory, UPR Arecibo. Since the Cobb-Douglas function is concave under certain constraints,
+the problem can be reduced to a convex optimization over the negative of the production function.
 
 \section{Cobb-Douglas Habitability Score}
+% Add the citations to Saha's papers.
 
-\subsection{Cobb-Douglas Habitability Production Function}
-The general form of the Cobb-Douglas production function is given by,
+\subsection{Cobb-Douglas Production Function}
+The general form of the Cobb-Douglas production function (CD-PF) is given by,
 \begin{equation}
@@
 
-% \begin{pointers}
-% \item Describe the Cobb Douglas Habitability Score.
-% \item Describe CRS, DRS and IRS\@.
-% \item Splitting the CDHS score into two values.
-% \item State the equations.
-% \item Score is estimated by maximizing the CDHPF fn.
-% \end{pointers}
+The CDHS is estimated by maximizing the CD-PF when the returns to scale are either decreasing or constant. This is
+marked by two constraints on the elasticity coefficients,
+\begin{itemize}
+\setlength\itemsep{0pt}
+\item When $\alpha+\beta < 1$, the CD-PF observes decreasing returns to scale.
+\item When $\alpha+\beta = 1$, the CD-PF observes constant returns to scale.
+\end{itemize}
+Further, if $0 < \alpha, \beta < 1$, the CD-PF is concave, and therefore has a maxima. The CD-PF can be adapted as a
+Habitability Production Function by setting the input parameters to the measurable parameters of an exoplanet and using
+these to estimate the elasticity coefficients.
+
+\subsection{Cobb-Douglas Habitability Score Estimation}
+Two types of Habitability Scores are calculated for each planet, the interior CDHS (CDHS\textsubscript{i}) and the
+surface CDHS (CDHS\textsubscript{s}), given by the following equations,
+\begin{subequations}
+  \begin{alignat}{4}
+    Y_1 &=\ {CDHS}_i\ &=\ &D^\alpha\ & \cdot\ &R^\beta\ &,\\
+    Y_2 &=\ {CDHS}_s\ &=\ &{V_e}^\gamma\ & \cdot\ &{T_s}^\delta\ &,
+  \end{alignat}
+\end{subequations}
+where, $D$, $R$, $V_e$ and $T_s$ are density, radius, escape velocity and surface temperature respectively. The final
+CDHS score is convex combination of the interior and surface CDHS values as given by,
+\begin{equation}
+  Y = w'\cdot Y_1 + w''\cdot Y_2,
+\end{equation}
+
+% Algorithm for CDHS estimation and explanation.
 

### 2017-12-01 docs/report/draft.tex
Some changes to report draft.

@@
 \usepackage{mathtools}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \newcommand{\pso}{Particle Swarm Optimization}
+
+% For algorithms
+\floatname{algorithm}{Procedure}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
 
@@
 \setlength\itemsep{0pt}
-\item When $\alpha+\beta < 1$, the CD-PF observes decreasing returns to scale.
-\item When $\alpha+\beta = 1$, the CD-PF observes constant returns to scale.
+\item When $\alpha+\beta < 1$, the CD-PF observes decreasing returns to scale (DRS).
+\item When $\alpha+\beta = 1$, the CD-PF observes constant returns to scale (CRS).
 \end{itemize}
@@
 
-% Algorithm for CDHS estimation and explanation.
+\begin{algorithm}
+  \begin{algorithmic}
+    \Require A dataset of exoplanets with density, radius, escape velocity and surface temperature in Earth Units.
+    \Ensure The $CDHS$ value for each exoplanet in the input dataset.
+
+    \State $CDHS \gets \{\}$
+    \ForAll{exoplanets $e$}
+      \State $D \gets$ $e.density$
+      \State $R \gets$ $e.radius$
+      \State $V_e \gets$ $e.escape\_velocity$
+      \State $T_s \gets$ $e.surface\_temperature$
+      \\ 
+      \State $Y_1 \gets \max\ D^\alpha\cdot R^\beta,\quad \text{subject to:}\quad \alpha+\beta<1;\quad 0<\alpha,\beta<1$
+      \State $Y_2 \gets \max\ {V_e}^\gamma\cdot {T_s}^\delta,\quad \text{subject to:}\quad \gamma+\delta<1;\quad
+      0<\gamma,\delta<1$
+      \State $Y.{DRS} \gets w'\cdot Y_1\ +\ w''\cdot Y_2$
+      \\
+      \State $Y_1 \gets \max\ D^\alpha\cdot R^\beta,\quad \text{subject to:}\quad \alpha+\beta=1;\quad 0<\alpha,\beta<1$
+      \State $Y_2 \gets \max\ {V_e}^\gamma\cdot {T_s}^\delta,\quad \text{subject to:}\quad \gamma+\delta=1;\quad
+      0<\gamma,\delta<1$
+      \State $Y.{CRS} \gets w'\cdot Y_1\ +\ w''\cdot Y_2$
+      \\
+      \State append $Y$ to $CDHS$
+    \EndFor
+  \end{algorithmic}
+  \caption{Estimating CDHS values.}
+  \label{estCDHS}
+\end{algorithm}
 

### 2018-01-11 source/ceesa/ceesa.py
Fixed a bug.

@@ -92,9 +92,10 @@ def evaluate_ceesa_values(exoplanets, fname='ceesa_{0}.csv', verbose=True,
                 continue
             score = np.round(ceesa(gbest), 4)
             if constraint == 'crs':
-                results.append([name, habc, *np.round(gbest, 4), 1, score, it])
+                results.append([name, habc, *np.round(gbest, 4), 1,
+                                score, it-99])
             else:
-                results.append([name, habc, *np.round(gbest, 4), score, it])
+                results.append([name, habc, *np.round(gbest, 4), score, it-99])
 
             if verbose:
                 print_results(_+1, total, results[-1])

### 2018-01-13 source/pso/pso.py
Removed a few comments.

@@ -1,10 +1,6 @@
 import numpy as np
 from numpy.random import uniform
 from scipy.spatial.distance import cdist
-import sys
-
-
-def dprint(*args, **kwargs): print(*args, **kwargs, end='\n\n')
 
 
 # Exception raised when Swarm does not converge.
@@ -79,7 +75,6 @@ def conmax_by_pso(fitness, start_points, constraints, friction=.8,
         # Update the local and global bests.
         position += velocity
         conmatrix = constraints(position)
-        #  print(conmatrix)
         to_update = (fitness(position) > fitness(lbest))
         to_update &= (conmatrix.sum(axis=1) < thresh)
 

### 2018-01-13 pso_test.py
Need to finish test case.

@@ -1,9 +1,22 @@
 #!/usr/bin/python
 
 import numpy as np
-from pso import Swarm
+from source import conmax_by_pso
 
-swarm = Swarm(10, ndim=1, fn=lambda x: -(x**2), friction=0.7,
-              learn_rate1=.5, learn_rate2=.5, max_velocity=4.0)
-swarm.converge()
-print(np.round(swarm.best_particle.position, 4))
+
+def fun(x):
+    return -((x.T[0]-2)**2 + (x.T[1]-1)**2)
+
+
+def constraints(x):
+    return np.apply_along_axis(lambda x: np.array((
+        np.max((x[0] - 2*x[1] + 1 - 1e-7, 0)),
+        np.max((-x[0] + 2*x[1] - 1 - 1e-7, 0)),
+        np.max((((x[0]**2)/4) + (x[1]**2) - 1), 0)
+    )), axis=1, arr=x)
+
+
+start = np.random.uniform(0, 2, (25, 2))
+gbest, it = conmax_by_pso(fun, start, constraints, learnrate1=.7,
+                          learnrate2=.7, max_velocity=1.)
+print(-1*fun(gbest))

### 2018-01-14 docs/report/draft.tex
Finished up to CO by PSO.

--- 2017-11-27 docs/report/draft.tex
@@
-\documentclass{article}
+\documentclass[10pt]{article}
 
+\usepackage[margin=1.2in]{geometry}
 \usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{enumitem}
 \usepackage{mathtools}
@@
 % Information for header.
-\title{Optimizing the Cobb Doughlas Habitability Production function using
-Particle Swarm Optimization}
+\title{Estimating Planetary Habitability via Particle Swarm Optimization of CES Production Functions.}
 \author{Abhijit J.\ Theophilus}
@@
 
+% For algorithms
+\floatname{algorithm}{Procedure}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+
+% For math
+\DeclareMathOperator*{\argmin}{arg\,min}
+
 \begin{document}
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 % Note to self, improve the flow of the report.
 % Find citations.
-Under most paradigms, optimization of a continuous function requires the calculation of a gradient. This might not
-always be feasible for non-polynomial functions in high-dimensional search spaces. This is further complicated when the
-optimization must be performed under constraints. \pso\ is a metaheuristic for continuous function optimization that does
-away with the need for a gradient. It employs a large number of particles that traverse the search space converging
-toward a global best solution encountered by at least one of the particles.
+The search for extra-terrestial life and potentially habitable extrasolar planets has been an international venture
+demanding large investments in cost and effort, since Frank Drake's attempt with Project Ozma in the mid-20th century.
+The first exoplanet was officially confirmed in 1992 which marked the start of a trend that has lasted 25 years and
+yielded over 3,700 confirmed exoplanets. There have been attempts to assess the habitability of these planets and to
+assign a score based on their similarity to Earth. Two such habitability scores are the Cobb-Douglas Habitability Score
+(CDHS) and the Constant Elasticity Earth Similarity Approach (CEESA) score. Estimating these scores involves maximizing
+a production function subjected to a set of constraints on its input variables.
 
-\pso\ is a distributed method that requires simple mathematical operators and few lines of code, making it a lucrative
-solution where computation resources are at a premium. Its implementation is highly parallelizable. It scales with the
-dimensionality of the search space. Constrained optimization can be incorporated within the standard algorithm through
-the use of a penalty function.
+Under most paradigms, maximizing a continuous function requires calculating a gradient. This may not always be feasible
+for non-polynomial functions in high-dimensional search spaces. Further, subjecting the input variables to constraints,
+as needed by CDHS and CEESA, are not always be straightforward to represent within the model. This paper presents an
+approach to Constrained Optimzation (CO) using the swarm intelligence metaheuristic. \pso\ (PSO) is a method for
+optimizing a continuous function that does away with the need for a gradient. It employs a large number of particles
+that traverse the search space converging toward a global best solution encountered by at least one of the particles.
+
+\pso\ is a distributed method that requires simple mathematical operators and short segments of code, making it a
+lucrative solution where computational resources are at a premium. Its implementation is highly parallelizable. It
+scales with the dimensionality of the search space. The standard PSO algorithm does not deal with constraints, but
+through variations in initializing and updating particles, constraints are straightforward to represent and adhere to,
+as seen in Section~\ref{subsec:copso}.
 
 % Pull out citations from the Applications paper.
-\pso\ has been successfully adapted to a wide range of design optimization problems, e.g., networks and VLSI design.
-\pso\ has also found applications in machine learning under clustering, feature detection and classification.  As a
-modelling paradigm, it has been used for constructing customer satisfaction models, MIDI music models, friction models,
-etc.
+PSO has been adapted to a wide range of design optimization problems, e.g., networks and VLSI design. It has found
+applications in machine learning under clustering, feature detection and classification. As a modeling paradigm, it has
+been used for constructing customer satisfaction models, modeling MIDI music and chaotic time series modeling.
 
-In this paper, \pso\ has been applied to estimate the Cobb Douglas Habitability Score of an exoplanet. The score is
-estimated by maximizing the Cobb-Douglas Habitability Production Function. The function considers four parameters,
-radius, mass, escape velocity and surface temperature, taken from the Exoplanets Catalog hosted by the Planetary
-Habitability Laboratory, UPR Arecibo. Since the Cobb-Douglas function is concave under certain constraints, the
-problem can be reduced to a convex optimization over the negative of the production function.
+This paper demonstrates the applicability of \pso\ in estimating CDHS and CEESA scores of an exoplanet by maximizing
+their respective production functions, discussed in Sections~\ref{subsec:cdhs} and~\ref{subsec:ceesa}. CDHS considers
+the planet's Radius, Mass, Escape Velocity and Surface Temperature, while CEESA includes a fifth parameter, the Orbital
+Eccentricity of the planet. The Exoplanets Catalog hosted by the Planetary Habitability Laboratory, UPR Arecibo records
+these parameters for each exoplanet in Earth Units. Section~\ref{sec:results} reports the performance of PSO and
+contrasts it against an earlier effort to estimate these scores using Stochastic Gradient Ascent (SGA).
 
-\section{Cobb-Douglas Habitability Score}
+\section{Habitability Scores}
+% Add the citations to Saha's papers.
 
-\subsection{Cobb-Douglas Habitability Production Function}
-The general form of the Cobb-Douglas production function is given by,
+\subsection{Cobb-Douglas Habitability Score}\label{subsec:cdhs}
+Estimating the Cobb-Douglas Habitability Score (CDHS) requires estimating an interior CDHS (CDHS\textsubscript{i}) and a
+surface CDHS (CDHS\textsubscript{s}) by maximizing the following production functions,
+\begin{subequations}
+  \begin{alignat}{4}
+    Y_i\ &=\ {CDHS}_i\ &=&\ R^\alpha.&D^\beta\label{eq:cdhsi}\\
+    Y_s\ &=\ {CDHS}_s\ &=&\ {V_e}^\gamma.&{T_s}^\delta\label{eq:cdhss}
+  \end{alignat}
+\end{subequations}
+where, $R$, $D$, $V_e$ and $T_s$ are density, radius, escape velocity and surface temperature respectively. $\alpha$,
+$\beta$, $\gamma$ and $\delta$ are the elasticity coefficients subject to,
 \begin{equation}
-  Y = k\cdot{(x_1)}^\alpha\cdot{(x_2)}^\beta,
+  0 < \alpha,\beta,\gamma,\delta < 1
 \end{equation}
-where $k$ is a constant that may be set according to requirement, $x_1$ and $x_2$ are the input parameters to the
-function, $Y$ is the total output production value, and $\alpha$ and $\beta$ are called the elasticity coefficients.
+Equations~\ref{eq:cdhsi} and~\ref{eq:cdhss} are convex when the returns to scale are either constant or decreasing,
+marked by two constraints on the elasticity coefficients,
+\begin{subequations}
+  \begin{align}
+    \alpha+\beta = 1,\quad\gamma+\delta = 1 && \text{for Constant Returns to Scale (CRS),}\\
+    \alpha+\beta < 1,\quad\gamma+\delta < 1 && \text{for Decreasing Returns to Scale (DRS).}
+  \end{align}
+\end{subequations}
+The final CDHS is the convex combination of the interior and surface CDHS values as given by,
+\begin{equation}
+  Y\ =\ w_i.Y_i + w_s.Y_s
+\end{equation}
 
-% \begin{pointers}
-% \item Describe the Cobb Douglas Habitability Score.
-% \item Describe CRS, DRS and IRS\@.
-% \item Splitting the CDHS score into two values.
-% \item State the equations.
-% \item Score is estimated by maximizing the CDHPF fn.
-% \end{pointers}
+\subsection{Constant Elasticity Earth Similarity Approach}\label{subsec:ceesa}
+The Constant Elasticity Earth Similarity Approach (CEESA) uses the following production function to estimate the
+habitability score of an exoplanet,
+\begin{equation}\label{eq:ceesa}
+  Y = {(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\eta/\rho}
+\end{equation}
+where, $E$ is the fifth parameter denoting Orbital Eccentricity. The value of $\rho$ lies within $0<\rho\leq 1$.
+The coefficients ($r$, $d$, $t$, $v$ and $e$) are constrained by,
+\begin{subequations}
+  \begin{align}
+      0 < r,d,t,v,e < 1\\
+      r+d+t+v+e = 1
+  \end{align}
+\end{subequations}
+The value of $\eta$ is constrained by the scale of production used,
+\begin{subequations}
+  \begin{align}
+    0 < \eta < 1 && \text{for Constant Returns to Scale (CRS),}\\
+    \eta = 1 && \text{for Decreasing Returns to Scale (DRS).}
+  \end{align}
+\end{subequations}
 
-\section{Particle Swarm Optimization}
-\begin{pointers}
-\item Equations. Modification for max.
-\item Constrained optimization modifications.
-\item Algorithm for PSO\@.
-\end{pointers}
+\section{Particle Swarm Optimization}\label{sec:pso}
+Particle Swarm Optimization (PSO) is a biologically inspired metaheuristic for finding the global minima of a function.
+Traditionally designed for unconstrained inputs, it works by iteratively converging a population of randomly initialized
+solutions, called particles, toward a globally optimal solution. Each particle in the population keeps track of its
+current position and the best solution it has encountered so far, called $pbest$. Each particle also has an associated
+randomized velocity used to traverse the search space. The swarm keeps track of the overall best solution, called
+$gbest$. Each iteration of the swarm updates the velocity of the particle towards its $pbest$ and the $gbest$ values.
+
+\subsection{PSO for Unconstrained Optimation}\label{subsec:uopso}
+Let $f(x)$ be the function to be minimized, where $x$ is a $d$-dimensional vector. $f(x)$ is also called the fitness
+function. Algorithm~\ref{alg:unop} outlines the approach to minimizing $f(x)$ using PSO.\@ A set of particles are randomly
+initialized with a position and a velocity, where $l$ and $u$ are the lower and upper boundaries of the search space.
+The position of the particle corresponds to its associated solution. The algorithm initializes each particle's $pbest$
+to its initial position. The $pbest$ position that corresponds to the minimum fitness is selected to be the $gbest$
+position of the swarm.
+
+On each iteration, the algorithm updates the velocity and position of each particle. For each particle, it picks two
+random numbers $u_g, u_p$ from a uniform distribution, $U(0,1)$ and updates the particle velocity as indicated in
+line~\ref{algline:vup}. Here, $\mu$ is the friction coefficient and $\lambda_g,\lambda_p$ are the global and particle
+learning rates. If the new position of the particle corresponds to a better fit than its $pbest$, the algorithm updates
+$pbest$ to the new position. Once the algorithm has updated all particles, it updates $gbest$ to the new overall best
+position. A suitable termination criteria for the swarm, under convex optimization, is  when the $gbest$ position has
+not changed by the end of the iteration.
+
+\begin{algorithm}
+  \begin{algorithmic}[1]
+    \Require{$f(x)$, the function to minimize.}
+    \Ensure{global minimum of $f(x)$.}
+    \For{each particle $i\gets 1,n$}
+      \State{$p_i \sim U(l,u)$}
+      \State{$v_i \sim U(-|u-l|, |u-l|)$}
+      \State{${pbest}_i \gets p_i$}
+    \EndFor\
+    \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
+    \Repeat\
+      \State{$oldbest \gets gbest$}
+      \For{each particle $i \gets 1,n$}
+        \State{$u_p, u_g \sim U(0,1)$}
+        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({gbest-p_i}) + \lambda_p.u_p.({{pbest}_i-p_i})$}\label{algline:vup}
+        \State{$p_i \gets p_i + v_i$}
+        \If{$f(p_i) < f({pbest}_i)$}
+          \State{${pbest}_i \gets p_i$}
+        \EndIf\
+      \EndFor\
+      \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
+    \Until{$|oldbest - gbest| < threshold$}
+    \State{\textbf{return} {$f(gbest)$}}
+  \end{algorithmic}
+  \caption{Algorithm for PSO.}\label{alg:unop}
+\end{algorithm}
+
+\subsection{PSO with Leaders for Constrained Optimization}\label{subsec:copso}
+Although PSO has eliminated the need to estimate the gradient of a function, as seen in Section~\ref{subsec:uopso}, it
+still is not suitable for constrained optimization. The standard PSO algorithm does not ensure that the initial
+solutions are feasible, and neither does it guarantee that the individual solutions will converge to a feasible global
+solution. Solving the initialization problem is straightforward, resample each random solution from the uniform
+distribution until every initial solution is feasible. To solve the convergence problem, each particle uses another
+particle's $pbest$ value, called $lbest$, instead of its own to update its velocity. Algorithm~\ref{alg:cop} describes
+this process.
+
+On each iteration, for each particle, the algorithm first picks two random numbers $u_g,u_p$ as before. It then selects
+a $pbest$ value from all particles in the swarm that is closest to the position of the particle being updated as its
+$lbest$. The $lbest$ value substitutes ${pbest}_i$ in the velocity update equation. While updating $pbest$ for the
+particle, the algorithm checks if the current fit is better than $pbest$, and performs the update if the current
+position satisfies all constraints.
+
+\begin{algorithm}
+  \begin{algorithmic}[1]
+    \Require{$f(x)$, the function to minimize.}
+    \Ensure{global minimum of $f(x)$.}
+    \For{each particle $i\gets 1,n$}
+      \Repeat
+        \State{$p_i \sim U(l,u)$}
+      \Until{$p_i$ satisfies all constraints}
+      \State{$v_i \sim U(-|u-l|, |u-l|)$}
+      \State{${pbest}_i \gets p_i$}
+    \EndFor\
+    \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
+    \Repeat\
+      \State{$oldbest \gets gbest$}
+      \For{each particle $i \gets 1,n$}
+        \State{$u_p, u_g \sim U(0,1)$}
+        \State{$lbest \gets \smashoperator{\argmin\limits_{{pbest}_j, j \in {1,n}}} {\|{pbest}_j -
+            p_i\|}^2$}\label{algline:lbest}
+        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({gbest-p_i}) + \lambda_p.u_p.({lbest-p_i})$}\label{algline:cvup}
+        \State{$p_i \gets p_i + v_i$}
+        \If{$f(p_i) < f({pbest}_i)$ \textbf{and} $p_i$ satisfies all constraints}
+          \State{${pbest}_i \gets p_i$}
+        \EndIf\
+      \EndFor\
+      \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
+    \Until{$|oldbest - gbest| < threshold$}
+    \State{\textbf{return} {$f(gbest)$}}
+  \end{algorithmic}
+  \caption{Algorithm for CO by PSO.}\label{alg:cop}
+\end{algorithm}
 

### 2018-01-15 docs/report/draft.tex
Finished the CO rep in draft.

@@
 \usepackage{mathtools}
+\usepackage{bm}
 
@@
 (CDHS) and the Constant Elasticity Earth Similarity Approach (CEESA) score. Estimating these scores involves maximizing
-a production function subjected to a set of constraints on its input variables.
+a production function while observing a set of constraints on the input variables.
 
@@
 for non-polynomial functions in high-dimensional search spaces. Further, subjecting the input variables to constraints,
-as needed by CDHS and CEESA, are not always be straightforward to represent within the model. This paper presents an
+as needed by CDHS and CEESA, are not always straightforward to represent within the model. This paper presents an
 approach to Constrained Optimzation (CO) using the swarm intelligence metaheuristic. \pso\ (PSO) is a method for
@@
 lucrative solution where computational resources are at a premium. Its implementation is highly parallelizable. It
-scales with the dimensionality of the search space. The standard PSO algorithm does not deal with constraints, but
-through variations in initializing and updating particles, constraints are straightforward to represent and adhere to,
-as seen in Section~\ref{subsec:copso}.
+scales with the dimensionality of the search space. The standard PSO algorithm does not deal with constraints but
+through variations in initializing and updating particles constraints are straightforward to represent and adhere to, as
+seen in Section~\ref{subsec:copso}.
 
 % Pull out citations from the Applications paper.
-PSO has been adapted to a wide range of design optimization problems, e.g., networks and VLSI design. It has found
+PSO has been adapted to a wide range of design optimization problems including network and VLSI design. It has found
 applications in machine learning under clustering, feature detection and classification. As a modeling paradigm, it has
@@
 \end{equation}
-Equations~\ref{eq:cdhsi} and~\ref{eq:cdhss} are convex when the returns to scale are either constant or decreasing,
-marked by two constraints on the elasticity coefficients,
+Equations~\ref{eq:cdhsi} and~\ref{eq:cdhss} are convex under either Constant Returns to Scale (CRS) or Decreasing
+Returns to Scale (DRS) marked by two constraints on the elasticity coefficients,
 \begin{subequations}
-  \begin{align}
-    \alpha+\beta = 1,\quad\gamma+\delta = 1 && \text{for Constant Returns to Scale (CRS),}\\
-    \alpha+\beta < 1,\quad\gamma+\delta < 1 && \text{for Decreasing Returns to Scale (DRS).}
-  \end{align}
+  \begin{alignat}{3}
+    \text{CRS:} & \quad\alpha+\beta = 1,&\quad\gamma+\delta = 1,\\
+    \text{DRS:} & \quad\alpha+\beta < 1,&\quad\gamma+\delta < 1.
+  \end{alignat}
 \end{subequations}
@@
 \begin{subequations}
-  \begin{align}
-    0 < \eta < 1 && \text{for Constant Returns to Scale (CRS),}\\
-    \eta = 1 && \text{for Decreasing Returns to Scale (DRS).}
-  \end{align}
+  \begin{alignat}{3}
+    \text{CRS:} & \quad 0 < \eta \leq 1,\\
+    \text{DRS:} & \quad \eta = 1.
+  \end{alignat}
 \end{subequations}
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
   \begin{algorithmic}[1]
@@
 particle, the algorithm checks if the current fit is better than $pbest$, and performs the update if the current
-position satisfies all constraints.
+position satisfies all constraints. The algorithm updates $gbest$ as before.
 
-\begin{algorithm}
+\begin{algorithm}[t]
   \begin{algorithmic}[1]
@@
     \For{each particle $i\gets 1,n$}
-      \Repeat
+      \Repeat\
         \State{$p_i \sim U(l,u)$}
@@
 \end{algorithm}
+
+\section{Representing the Constrained Optimization Problem}\label{sec:rep}
+A Constrained Optimization problem can represented as,
+\begin{equation*}
+  \begin{aligned}
+    & \underset{x}{\text{minimize}}
+    & & f(x) \\
+    & \text{subject to}
+    & & g_k(x) \leq 0,&\; k = 1,\dots,q\\
+    &&& h_l(x) = 0,&\quad l = 1,\dots,r.
+  \end{aligned}
+\end{equation*}
+
+Introducing an error threshold $\epsilon$ can convert strict inequalities of the form ${g_k}^{\prime}(x) < 0$ to
+nonstrict inequalities of the form $g_k(x) = {g_k}^{\prime}(x) + \epsilon \leq 0$. Adding a tolerance $\tau$ transforms
+equality constraints to a pair of inequalities,
+\begin{equation*}
+  \begin{aligned}
+    g_{(q+l)}(x) &= h_l(x) - \tau &\leq 0,&\quad l = 1,\dots,r\\
+    g_{(q+r+l)}(x) &= {-}h_l(x) - \tau &\leq 0,&\quad l = 1,\dots,r
+  \end{aligned}
+\end{equation*}
+Thus, $r$ equality constraints become $2r$ inequality constraints, raising the total number of constraints, denoted by
+$s$, to $s = q + 2r$. For each solution $p_i$, $\bm{c_i}$ denotes the constraint vector where, $c_{ik} = \max\{g_k(p_i),
+0\},~k=1,\dots,s$. When $\bm{c_i} = \bm{0}$, the solution $p_i$ lies within the feasible region. When $c_{ik} > 0$, the
+solution $p_i$ violates the $k$\textsuperscript{th} constraint.
+
+Following these guidelines to represent a CO problem, the CDHS estimation under CRS becomes,
+\begin{center}
+\setlength{\tabcolsep}{30pt}
+\begin{tabular}{l | l}
+\begin{equation*}
+  \begin{aligned}
+    & \underset{\alpha,\beta}{\text{minimize:}}
+    &  Y_i = {-}R^\alpha.D^\beta,\\
+    &  \text{subject to:}
+    &  {-}\alpha + \epsilon \leq 0,\\
+    && \alpha - 1 + \epsilon \leq 0,\\
+    && {-}\beta + \epsilon  \leq 0,\\
+    && \beta - 1 + \epsilon \leq 0,\\
+    && \alpha + \beta - 1 - \tau \leq 0,\\
+    && {-}(\alpha + \beta - 1) - \tau \leq 0.
+  \end{aligned}
+\end{equation*}
+&
+\begin{equation*}
+  \begin{aligned}
+    &  \underset{\gamma,\delta}{\text{minimize:}}
+    &  Y_s = {-}{V_e}^\gamma.{T_s}^\delta\\
+    &  \text{subject to:}
+    &  {-}\gamma + \epsilon \leq 0,\\
+    && \gamma - 1 + \epsilon \leq 0,\\
+    && {-}\delta + \epsilon  \leq 0,\\
+    && \delta - 1 + \epsilon \leq 0,\\
+    && \gamma + \delta - 1 - \tau \leq 0,\\
+    && {-}(\gamma + \delta - 1) - \tau \leq 0,
+  \end{aligned}
+\end{equation*}
+\end{tabular}
+\end{center}
+while for DRS, the last two constraints for $Y_i$ and $Y_s$ are replaced with,
+\begin{center}
+\begin{tabular}{l c r}
+\begin{equation*}
+  \alpha + \beta + \epsilon - 1 \leq 0
+\end{equation*}
+&&
+\begin{equation*}
+  \gamma + \delta + \epsilon - 1 \leq 0
+\end{equation*}
+\end{tabular}
+\end{center}
+
+The CEESA score estimation for DRS is represented as,
+\begin{equation*}
+  \begin{aligned}
+    & \underset{r,d,t,v,e,\rho,\eta}{\text{minimize}}
+    & Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\eta/\rho}\\
+    & \text{subject to}
+    &  {-}\phi + \epsilon \leq 0, &\quad\quad \forall \phi\in\{r,d,t,v,e,\rho,\eta\}\\
+    && \phi - 1 + \epsilon \leq 0, &\quad\quad \forall \phi\in\{r,d,t,v,e,\eta\}\\
+    && \rho - 1 \leq 0,\\
+    && (r+d+t+v+e-1) - \tau \leq 0,\\
+    && (1-r-d-t-v-e) - \tau \leq 0,
+  \end{aligned}
+\end{equation*}
+while for CRS, there is no need for the parameter $\eta$. The objective function for the problem becomes,
+\begin{equation*}
+  \underset{r,d,t,v,e,\rho}{\text{minimize}}\quad
+  Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{1/\rho}
+\end{equation*}
 

### 2018-01-16 docs/report/draft.tex
Only results and conclusion pending.

@@
 \usepackage{algpseudocode}
+\usepackage{booktabs}
 \usepackage{enumitem}
@@
 \usepackage{bm}
+\usepackage{siunitx}
+
+\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
 
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\sgn}{sgn}
+\DeclareMathOperator*{\pbest}{\mathit{pbest}}
+\DeclareMathOperator*{\gbest}{\mathit{gbest}}
+\DeclareMathOperator*{\oldbest}{\mathit{oldbest}}
+\DeclareMathOperator*{\lbest}{\mathit{lbest}}
 
@@
 \section{Introduction}\label{sec:intro}
-% Note to self, improve the flow of the report.
-% Find citations.
-The search for extra-terrestial life and potentially habitable extrasolar planets has been an international venture
+
+The search for extra-terrestrial life and potentially habitable extrasolar planets has been an international venture
 demanding large investments in cost and effort, since Frank Drake's attempt with Project Ozma in the mid-20th century.
@@
 as needed by CDHS and CEESA, are not always straightforward to represent within the model. This paper presents an
-approach to Constrained Optimzation (CO) using the swarm intelligence metaheuristic. \pso\ (PSO) is a method for
+approach to Constrained Optimization (CO) using the swarm intelligence metaheuristic. \pso\ (PSO) is a method for
 optimizing a continuous function that does away with the need for a gradient. It employs a large number of particles
@@
 \begin{equation}\label{eq:ceesa}
-  Y = {(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\eta/\rho}
+  Y = {(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\frac{\eta}{\rho}}
 \end{equation}
@@
 \section{Particle Swarm Optimization}\label{sec:pso}
+
 Particle Swarm Optimization (PSO) is a biologically inspired metaheuristic for finding the global minima of a function.
@@
 solutions, called particles, toward a globally optimal solution. Each particle in the population keeps track of its
-current position and the best solution it has encountered so far, called $pbest$. Each particle also has an associated
+current position and the best solution it has encountered so far, called $\pbest$. Each particle also has an associated
 randomized velocity used to traverse the search space. The swarm keeps track of the overall best solution, called
-$gbest$. Each iteration of the swarm updates the velocity of the particle towards its $pbest$ and the $gbest$ values.
+$\gbest$. Each iteration of the swarm updates the velocity of the particle towards its $\pbest$ and the $\gbest$ values.
 
-\subsection{PSO for Unconstrained Optimation}\label{subsec:uopso}
+\subsection{PSO for Unconstrained Optimization}\label{subsec:uopso}
 Let $f(x)$ be the function to be minimized, where $x$ is a $d$-dimensional vector. $f(x)$ is also called the fitness
@@
 initialized with a position and a velocity, where $l$ and $u$ are the lower and upper boundaries of the search space.
-The position of the particle corresponds to its associated solution. The algorithm initializes each particle's $pbest$
-to its initial position. The $pbest$ position that corresponds to the minimum fitness is selected to be the $gbest$
+The position of the particle corresponds to its associated solution. The algorithm initializes each particle's $\pbest$
+to its initial position. The $\pbest$ position that corresponds to the minimum fitness is selected to be the $\gbest$
 position of the swarm.
@@
 line~\ref{algline:vup}. Here, $\mu$ is the friction coefficient and $\lambda_g,\lambda_p$ are the global and particle
-learning rates. If the new position of the particle corresponds to a better fit than its $pbest$, the algorithm updates
-$pbest$ to the new position. Once the algorithm has updated all particles, it updates $gbest$ to the new overall best
-position. A suitable termination criteria for the swarm, under convex optimization, is  when the $gbest$ position has
+learning rates. If the new position of the particle corresponds to a better fit than its $\pbest$, the algorithm updates
+$\pbest$ to the new position. Once the algorithm has updated all particles, it updates $\gbest$ to the new overall best
+position. A suitable termination criteria for the swarm, under convex optimization, is  when the $\gbest$ position has
 not changed by the end of the iteration.
@@
     \For{each particle $i\gets 1,n$}
-      \State{$p_i \sim U(l,u)$}
-      \State{$v_i \sim U(-|u-l|, |u-l|)$}
-      \State{${pbest}_i \gets p_i$}
+    \State{$p_i \sim {U(l,u)}^d$}
+    \State{$v_i \sim {U(-|u-l|, |u-l|)}^d$}
+      \State{${\pbest}_i \gets p_i$}
     \EndFor\
-    \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
+    \State{$\gbest \gets \smashoperator{\argmin\limits_{{\pbest}_i,\,i=1\dots n}} f({\pbest}_i)$}
     \Repeat\
-      \State{$oldbest \gets gbest$}
-      \For{each particle $i \gets 1,n$}
+      \State{$\oldbest \gets \gbest$}
+      \For{each particle $i \gets 1\dots n$}
         \State{$u_p, u_g \sim U(0,1)$}
-        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({gbest-p_i}) + \lambda_p.u_p.({{pbest}_i-p_i})$}\label{algline:vup}
+        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({\gbest-p_i}) + \lambda_p.u_p.({{\pbest}_i-p_i})$}\label{algline:vup}
         \State{$p_i \gets p_i + v_i$}
-        \If{$f(p_i) < f({pbest}_i)$}
-          \State{${pbest}_i \gets p_i$}
+        \If{$f(p_i) < f({\pbest}_i)$}
+          \State{${\pbest}_i \gets p_i$}
         \EndIf\
       \EndFor\
-      \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
-    \Until{$|oldbest - gbest| < threshold$}
-    \State{\textbf{return} {$f(gbest)$}}
+      \State{$\gbest \gets \smashoperator{\argmin\limits_{{\pbest}_i,\,i=1\dots n}} f({\pbest}_i)$}
+    \Until{$|\oldbest - \gbest| < threshold$}
+    \State{\textbf{return} {$f(\gbest)$}}
   \end{algorithmic}
@@
 distribution until every initial solution is feasible. To solve the convergence problem, each particle uses another
-particle's $pbest$ value, called $lbest$, instead of its own to update its velocity. Algorithm~\ref{alg:cop} describes
+particle's $\pbest$ value, called $\lbest$, instead of its own to update its velocity. Algorithm~\ref{alg:cop} describes
 this process.
@@
 On each iteration, for each particle, the algorithm first picks two random numbers $u_g,u_p$ as before. It then selects
-a $pbest$ value from all particles in the swarm that is closest to the position of the particle being updated as its
-$lbest$. The $lbest$ value substitutes ${pbest}_i$ in the velocity update equation. While updating $pbest$ for the
-particle, the algorithm checks if the current fit is better than $pbest$, and performs the update if the current
-position satisfies all constraints. The algorithm updates $gbest$ as before.
+a $\pbest$ value from all particles in the swarm that is closest to the position of the particle being updated as its
+$\lbest$. The $\lbest$ value substitutes ${\pbest}_i$ in the velocity update equation. While updating $\pbest$ for the
+particle, the algorithm checks if the current fit is better than $\pbest$, and performs the update if the current
+position satisfies all constraints. The algorithm updates $\gbest$ as before.
 
@@
       \Repeat\
-        \State{$p_i \sim U(l,u)$}
+      \State{$p_i \sim {U(l,u)}^d$}
       \Until{$p_i$ satisfies all constraints}
-      \State{$v_i \sim U(-|u-l|, |u-l|)$}
-      \State{${pbest}_i \gets p_i$}
+      \State{$v_i \sim {U(-|u-l|, |u-l|)}^d$}
+      \State{${\pbest}_i \gets p_i$}
     \EndFor\
-    \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
+    \State{$\gbest \gets \smashoperator{\argmin\limits_{{\pbest}_i,\,i=1\dots n}} f({\pbest}_i)$}
     \Repeat\
-      \State{$oldbest \gets gbest$}
-      \For{each particle $i \gets 1,n$}
+      \State{$\oldbest \gets \gbest$}
+      \For{each particle $i \gets 1\dots n$}
         \State{$u_p, u_g \sim U(0,1)$}
-        \State{$lbest \gets \smashoperator{\argmin\limits_{{pbest}_j, j \in {1,n}}} {\|{pbest}_j -
+        \State{$\lbest \gets \smashoperator{\argmin\limits_{{\pbest}_j,\,j=1\dots n}} {\|{\pbest}_j -
             p_i\|}^2$}\label{algline:lbest}
-        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({gbest-p_i}) + \lambda_p.u_p.({lbest-p_i})$}\label{algline:cvup}
+        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({\gbest-p_i}) + \lambda_p.u_p.({\lbest-p_i})$}\label{algline:cvup}
         \State{$p_i \gets p_i + v_i$}
-        \If{$f(p_i) < f({pbest}_i)$ \textbf{and} $p_i$ satisfies all constraints}
-          \State{${pbest}_i \gets p_i$}
+        \If{$f(p_i) < f({\pbest}_i)$ \textbf{and} $p_i$ satisfies all constraints}
+          \State{${\pbest}_i \gets p_i$}
         \EndIf\
       \EndFor\
-      \State{$gbest \gets \smashoperator{\argmin\limits_{{pbest}_i, i \in {1,n}}} f({pbest}_i)$}
-    \Until{$|oldbest - gbest| < threshold$}
-    \State{\textbf{return} {$f(gbest)$}}
+      \State{$\gbest \gets \smashoperator{\argmin\limits_{{\pbest}_i,\,i=1\dots n}} f({\pbest}_i)$}
+    \Until{$|\oldbest - \gbest| < threshold$}
+    \State{\textbf{return} {$f(\gbest)$}}
   \end{algorithmic}
@@
 
-\section{Representing the Constrained Optimization Problem}\label{sec:rep}
+\section{Representing the Problem}\label{sec:rep}
+
 A Constrained Optimization problem can represented as,
@@
     & \text{subject to}
-    & & g_k(x) \leq 0,&\; k = 1,\dots,q\\
-    &&& h_l(x) = 0,&\quad l = 1,\dots,r.
+    & & g_k(x) \leq 0,&\; k = 1\dots q,\\
+    &&& h_l(x) = 0,&\quad l = 1\dots r.
   \end{aligned}
@@
 Introducing an error threshold $\epsilon$ can convert strict inequalities of the form ${g_k}^{\prime}(x) < 0$ to
-nonstrict inequalities of the form $g_k(x) = {g_k}^{\prime}(x) + \epsilon \leq 0$. Adding a tolerance $\tau$ transforms
+non-strict inequalities of the form $g_k(x) = {g_k}^{\prime}(x) + \epsilon \leq 0$. Adding a tolerance $\tau$ transforms
 equality constraints to a pair of inequalities,
@@
   \begin{aligned}
-    g_{(q+l)}(x) &= h_l(x) - \tau &\leq 0,&\quad l = 1,\dots,r\\
-    g_{(q+r+l)}(x) &= {-}h_l(x) - \tau &\leq 0,&\quad l = 1,\dots,r
+    g_{(q+l)}(x) &= h_l(x) - \tau &\leq 0,&\quad l = 1\dots r,\\
+    g_{(q+r+l)}(x) &= {-}h_l(x) - \tau &\leq 0,&\quad l = 1\dots r.
   \end{aligned}
@@
 Thus, $r$ equality constraints become $2r$ inequality constraints, raising the total number of constraints, denoted by
-$s$, to $s = q + 2r$. For each solution $p_i$, $\bm{c_i}$ denotes the constraint vector where, $c_{ik} = \max\{g_k(p_i),
-0\},~k=1,\dots,s$. When $\bm{c_i} = \bm{0}$, the solution $p_i$ lies within the feasible region. When $c_{ik} > 0$, the
-solution $p_i$ violates the $k$\textsuperscript{th} constraint.
+$s$, to $s = q + 2r$. For each solution $p_i$, $c_i$ denotes the constraint vector where, $c_{ik} = \max\{g_k(p_i),
+0\},~k=1\dots s$. When $c_{ik} = 0,~\forall k=1\dots s$, the solution $p_i$ lies within the feasible region. When $c_{ik} > 0$,
+the solution $p_i$ violates the $k$\textsuperscript{th} constraint.
 
-Following these guidelines to represent a CO problem, the CDHS estimation under CRS becomes,
+Following these guidelines to represent a CO problem, CDHS estimation under CRS becomes,
+\begin{equation}\label{eq:cdhscrs}
+  \begin{aligned}
+    & \underset{\alpha,\beta,\gamma,\delta}{\text{minimize:}}
+    & Y_i = {-}R^\alpha.D^\beta,\ Y_s = {-}{V_e}^\gamma.{T_s}^\delta\\
+    & \text{subject to:}
+    &  {-}\phi + \epsilon \leq 0,\quad\phi - 1 + \epsilon \leq 0, &\quad\quad \forall \phi\in\{\alpha,\beta,\gamma,\delta\}\\
+    && (\alpha+\beta-1) - \tau \leq 0,\quad(\gamma+\delta-1) - \tau \leq 0,\\
+    && (1-\alpha-\beta) - \tau \leq 0,\quad(1-\gamma-\delta) - \tau \leq 0,
+  \end{aligned}
+\end{equation}
+but with DRS the last two constraints for $Y_i$ and $Y_s$ are replaced with,
+\begin{equation}\label{eq:cdhsdrs}
+  \begin{aligned}
+    \alpha + \beta + \epsilon - 1 &\leq 0,\\
+    \gamma + \delta + \epsilon - 1 &\leq 0.
+  \end{aligned}
+\end{equation}
+
+The CEESA score estimation for DRS is represented as,
+\begin{equation}\label{eq:ceesadrs}
+  \begin{aligned}
+    & \underset{r,d,t,v,e,\rho,\eta}{\text{minimize}}
+    & Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\frac{\eta}{\rho}}\\
+    & \text{subject to}
+    &   {-}\phi + \epsilon \leq 0,\quad\phi - 1 + \epsilon \leq 0, &\quad\quad \forall \phi\in\{r,d,t,v,e,\eta\}\\
+    && \rho - 1 \leq 0,\quad \rho - 1 + \epsilon \leq 0,\\
+    && (r+d+t+v+e-1) - \tau \leq 0,\\
+    && (1-r-d-t-v-e) - \tau \leq 0.
+  \end{aligned}
+\end{equation}
+Under CRS there is no need for the parameter $\eta$. The objective function for the problem becomes,
+\begin{equation}\label{eq:ceesacrs}
+  \underset{r,d,t,v,e,\rho}{\text{minimize}}\quad
+  Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\frac{1}{\rho}}.
+\end{equation}
+
+\section{Experiment and Results}
+
+The data set used for analysis is the Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory
+(PHL). The catalog records observed and modeled parameters for exoplanets confirmed by the Extrasolar Planets
+Encyclopedia. Table~\ref{tab:param} describes the parameters from the PHL Exoplanets Catalog (PHL-EC) used for the
+experiment. Since surface temperature and eccentricity are not recorded in Earth Units, we normalized these values by
+dividing them with Earth's surface temperature (288K) and eccentricity (0.017). PHL-EC assumes an Eccentricity of 0 when
+unavailable.
+
+\begin{table}
   \begin{center}
-\setlength{\tabcolsep}{30pt}
-\begin{tabular}{l | l}
-\begin{equation*}
-  \begin{aligned}
-    & \underset{\alpha,\beta}{\text{minimize:}}
-    &  Y_i = {-}R^\alpha.D^\beta,\\
-    &  \text{subject to:}
-    &  {-}\alpha + \epsilon \leq 0,\\
-    && \alpha - 1 + \epsilon \leq 0,\\
-    && {-}\beta + \epsilon  \leq 0,\\
-    && \beta - 1 + \epsilon \leq 0,\\
-    && \alpha + \beta - 1 - \tau \leq 0,\\
-    && {-}(\alpha + \beta - 1) - \tau \leq 0.
-  \end{aligned}
-\end{equation*}
-&
-\begin{equation*}
-  \begin{aligned}
-    &  \underset{\gamma,\delta}{\text{minimize:}}
-    &  Y_s = {-}{V_e}^\gamma.{T_s}^\delta\\
-    &  \text{subject to:}
-    &  {-}\gamma + \epsilon \leq 0,\\
-    && \gamma - 1 + \epsilon \leq 0,\\
-    && {-}\delta + \epsilon  \leq 0,\\
-    && \delta - 1 + \epsilon \leq 0,\\
-    && \gamma + \delta - 1 - \tau \leq 0,\\
-    && {-}(\gamma + \delta - 1) - \tau \leq 0,
-  \end{aligned}
-\end{equation*}
+    \begin{tabular}{l l l}
+      \toprule
+      \textbf{Parameter} & \textbf{Description} & \textbf{Unit}\\
+      \midrule
+      P. Radius       & Estimated radius      & Earth Units (EU)\\
+      P. Density      & Density               & Earth Units (EU)\\
+      P. Esc Vel      & Escape velocity       & Earth Units (EU)\\
+      P. Ts Mean      & Surface temperature   & Kelvin (K)\\
+      P. Eccentricity & Orbital eccentricity\\
+      \bottomrule
     \end{tabular}
   \end{center}
-while for DRS, the last two constraints for $Y_i$ and $Y_s$ are replaced with,
-\begin{center}
-\begin{tabular}{l c r}
-\begin{equation*}
-  \alpha + \beta + \epsilon - 1 \leq 0
-\end{equation*}
-&&
-\begin{equation*}
-  \gamma + \delta + \epsilon - 1 \leq 0
-\end{equation*}
-\end{tabular}
-\end{center}
+  \caption{Parameters from the PHL-EC used for the experiment.}\label{tab:param}
+\end{table}
 
-The CEESA score estimation for DRS is represented as,
-\begin{equation*}
-  \begin{aligned}
-    & \underset{r,d,t,v,e,\rho,\eta}{\text{minimize}}
-    & Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\eta/\rho}\\
-    & \text{subject to}
-    &  {-}\phi + \epsilon \leq 0, &\quad\quad \forall \phi\in\{r,d,t,v,e,\rho,\eta\}\\
-    && \phi - 1 + \epsilon \leq 0, &\quad\quad \forall \phi\in\{r,d,t,v,e,\eta\}\\
-    && \rho - 1 \leq 0,\\
-    && (r+d+t+v+e-1) - \tau \leq 0,\\
-    && (1-r-d-t-v-e) - \tau \leq 0,
-  \end{aligned}
-\end{equation*}
-while for CRS, there is no need for the parameter $\eta$. The objective function for the problem becomes,
-\begin{equation*}
-  \underset{r,d,t,v,e,\rho}{\text{minimize}}\quad
-  Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{1/\rho}
-\end{equation*}
+The implemenation uses $n=25$ particles to traverse the search space with learning rates of $\lambda_g=0.8$ and
+$\lambda_p=0.2$. Velocity is regulated through a friction coefficient of $\mu=0.6$ and bound by $\pm1.0$. We have used
+an error threshold of $\epsilon=\num{1e-6}$ in converting strict inequalities to non-strict inequalities, and a
+tolerance of $\tau=\num{1e-7}$ when transforming an equality constraint to a pair of inequalities. Further
+implementation details are discussed in Appendix~\ref{app:imp}.
 
-\section{Experiment}
-\begin{pointers}
-\item Discussing the data set.
-\item Parameters for the constrained optimization.
-\item Implementation method.
-\item Ensuring convergence.
-\end{pointers}
+Tables~\ref{tab:cdhscrs} and~\ref{tab:cdhsdrs} record the difference in the CDHS values as estimated by \pso\ and
+Stochastic Gradient Ascent for a sample of planets under CRS and DRS respectively. Tables~\ref{tab:ceesacrs}
+and~\ref{tab:ceesadrs} record the same for the CEESA scores.
+
+% GJ 667C e
+% GJ 667C f
+% GJ 832 c
+% HD 40307 g
+% Kapteyn’s b
+% Kepler-61 b
+% Kepler-62 e
+% Kepler-62 f
+% Kepler-174 d
+% Kepler-186 f
+% Kepler-283 c
+% Kepler-296 f
+% GJ 667C c
+% GJ 163 c
 
@@
 
+\appendix
+
+\section{Improving Performance}\label{app:imp}
+
+Matrices $P$ and $V$ represent current position and velocity, where the $i$\textsuperscript{th} row of each correspond
+to the position and velocity of particle $i$. Each row of the matrix $L$ is the leader for the particle in the
+corresponding row of $P$. The constraint matrix $C$ is constructed by stacking the constraint vectors $c_i$ described in
+Section~\ref{sec:rep}. Let $r',r''$ be two random vectors of length $n$ drawn from the uniform distribution
+${U(0,1)}^n$. Let $X_i$ denote the $i$\textsuperscript{th} row of matrix $X$. The implementation of each iteration while
+updating particles in Algorithm~\ref{alg:cop} reduces to,
+
+\begin{align*}
+  g' &= f([\gbest])\\
+  L  &= {[\argmin\limits_{pbest_j,\ j = 1\dots n} {\|pbest_j - P_i\|}^2\ \mid\ \forall i = 1\dots n\quad ]}^T\\
+  V  &= \mu.V + \lambda_g\begin{bmatrix}
+    {r_1}'(\gbest - P_1)\\
+    {r_2}'(\gbest - P_2)\\
+    \vdots\\
+    {r_n}'(\gbest - P_n)
+  \end{bmatrix} + \lambda_p\begin{bmatrix}
+    {r_1}''(L_1 - P_1)\\
+    {r_2}''(L_2 - P_2)\\
+    \vdots\\
+    {r_n}''(L_n - P_n)
+  \end{bmatrix}\\
+  V  &= [\quad\sgn(v_{ij})*\max\{|v_{max}|,|v_{ij}|\}\ \mid\ \forall i=1\dots n,\forall j=1\dots d\quad]\\
+  P  &= P + V\\
+\end{align*}
+
 \end{document}

### 2018-01-16 source/cdhs/cdhs.py

@@ -69,8 +69,6 @@ def evaluate_cdhs_values(exoplanets, fname='cdhs_{0}.csv', verbose=True,
         kwargs:
             The parameters for the Swarm.
     """
-    exoplanets.dropna(how='any', inplace=True)
-    exoplanets.reset_index(drop=True, inplace=True)
     total = len(exoplanets)
 
     for constraint in ('crs', 'drs'):

### 2018-01-16 source/ceesa/ceesa.py

@@ -66,8 +66,6 @@ def evaluate_ceesa_values(exoplanets, fname='ceesa_{0}.csv', verbose=True,
         kwargs:
             The parameters for the Swarm.
     """
-    exoplanets.fillna(value=0, inplace=True)
-    exoplanets.reset_index(drop=True, inplace=True)
     total = len(exoplanets)
 
     for constraint in ('crs', 'drs'):

### 2018-01-16 source/exoplanets/load_data.py

@@ -27,7 +27,8 @@ exoplanets.rename(columns={
     'P. Eccentricity': 'Eccentricity',
     'P. Habitable Class': 'Habitable',
 }, inplace=True)
-exoplanets.dropna(how='all', inplace=True)
+exoplanets.dropna(how='any', inplace=True)
+exoplanets.reset_index(drop=True, inplace=True)
 exoplanets['STemp'] /= 288
 exoplanets['Eccentricity'] /= 0.017
 exoplanets = _round(exoplanets)

### 2018-01-16 docs/report/draft.tex
Iterations are stored for CDHS.

@@
 
-Tables~\ref{tab:cdhscrs} and~\ref{tab:cdhsdrs} record the difference in the CDHS values as estimated by \pso\ and
-Stochastic Gradient Ascent for a sample of planets under CRS and DRS respectively. Tables~\ref{tab:ceesacrs}
-and~\ref{tab:ceesadrs} record the same for the CEESA scores.
+Tables~\ref{tab:cdhscrs} and~\ref{tab:cdhsdrs} record the CDHS values for a sample of exoplanets under CRS and DRS
+respectively. Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the CEESA scores. All tables also
+record the number of iterations taken to converge to a stable $\gbest$ value.
 

### 2018-01-16 source/cdhs/cdhs.py

@@ -9,13 +9,15 @@ from ..pso import conmax_by_pso, SwarmConvergeError
 
 
 # Miscellaneous Consts.
-MESSAGE = '{:25}{:8.4f}{:8.4f}{:10.4f}{:8.4f}{:8.4f}{:10.4f}{:10.4f}{:>7.5}'
-TITLE = '{:25}{:>8}{:>8}{:>10}{:>8}{:>8}{:>10}{:>10}{:>7.5}'
-ERROR = '{:25}{:^69}'
-HEADERS = ('Name', 'A', 'B', 'CDHSi', 'G', 'D', 'CDHSs', 'CDHS', 'Cls')
+MESSAGE = '{:25}{:>7.5}{:8.4f}{:8.4f}{:10.4f}'\
+          '{:8.4f}{:8.4f}{:10.4f}{:10.4f}{:5}{:5}'
+TITLE = '{:25}{:>7.5}{:>8}{:>8}{:>10}{:>8}{:>8}{:>10}{:>10}{:>5}{:>5}'
+ERROR = '{:25}{:^79}'
+HEADERS = ('Name', 'Cls', 'A', 'B', 'CDHSi',
+           'G', 'D', 'CDHSs', 'CDHS', 'IterI', 'IterS')
 ERR_CDHSi = '** CDHSi convergence failed. **'
 ERR_CDHSs = '** CDHSs convergence failed. **'
-TOTAL_CHAR = 94
+TOTAL_CHAR = 104
 PROGRESS_BAR = '[{:' + str(TOTAL_CHAR - 10) + '}]  ({:>3}%)'
 
 
@@ -23,13 +25,9 @@ PROGRESS_BAR = '[{:' + str(TOTAL_CHAR - 10) + '}]  ({:>3}%)'
 def print_header(constraint, headers):
     """Print the header of the output table."""
     spaces = (TOTAL_CHAR//2 - len(constraint)//2) * ' '
-
     print('\n' + spaces + constraint.upper())
     print(spaces + '-'*len(constraint) + '\n')
-    print(TITLE.format(*headers))
-
-    print('-' * TOTAL_CHAR, end='\t\t')
-    print('  inn : sur  ')
+    print(TITLE.format(*headers), '-' * TOTAL_CHAR, sep='\n')
 
 
 def print_error(name, err):
@@ -37,11 +35,9 @@ def print_error(name, err):
     print(ERROR.format(name, err))
 
 
-def print_results(it, total, values, iters):
+def print_results(it, total, values):
[…]

### 2018-01-16 docs/report/draft.tex
Filling in tables.

@@
 Tables~\ref{tab:cdhscrs} and~\ref{tab:cdhsdrs} record the CDHS values for a sample of exoplanets under CRS and DRS
-respectively. Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the CEESA scores. All tables also
+respectively at $w_i=0.99$ and $w_s=0.01$. Each of these tables record the parameters of Equations~\ref{eq:cdhscrs}
+and~\ref{eq:cdhsdrs} and the objective function values. Columns $i_i$ and $i_s$ specify the number of iterations taken
+to converge to a stable $\gbest$ value. Under the Class column there are four categories for the planets ---
+Psychroplanets (psy), Mesoplanets (mes), Non-Habitable planets (non) and Hypopsychroplanets (hyp).
+
+Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the CEESA scores. All tables also
 record the number of iterations taken to converge to a stable $\gbest$ value.
 
-% GJ 667C e
-% GJ 667C f
-% GJ 832 c
-% HD 40307 g
-% Kapteyn’s b
-% Kepler-61 b
-% Kepler-62 e
-% Kepler-62 f
-% Kepler-174 d
-% Kepler-186 f
-% Kepler-283 c
-% Kepler-296 f
-% GJ 667C c
-% GJ 163 c
+\begin{table}
+  \centering
+  \begin{tabular}{l r r r r r r r r r r}
+    \toprule
+    Name & Class & $\alpha$ & $\beta$ & $\mathit{Y}_i$ & ${i}_i$ & 
+    $\gamma$ & $\delta$ & $\mathit{Y}_s$ & ${i}_s$ & $\mathit{CDHS}$\\
+    \midrule
+    GJ 176 b & non & 0.46 & 0.54 & 1.90 & 50 & 0.11 & 0.89 & 2.11 & 61 & 1.90\\
+    GJ 667 C b & non & 0.42 & 0.58 & 1.71 & 58 & 0.69 & 0.31 & 1.81 & 54 & 1.71\\
+    GJ 667 C e & psy & 0.13 & 0.87 & 1.40 & 50 & 0.26 & 0.74 & 1.39 & 55 & 1.40\\
+    GJ 667 C f & psy & 0.53 & 0.47 & 1.40 & 48 & 0.87 & 0.13 & 1.39 & 47 & 1.40\\
+    GJ 3634 b & non & 0.41 & 0.59 & 1.89 & 58 & 0.72 & 0.28 & 2.09 & 48 & 1.89\\
+    HD 20794 c & non & 0.26 & 0.74 & 1.35 & 50 & 0.10 & 0.90 & 1.34 & 58 & 1.35\\
+    HD 40307 e & non & 0.17 & 0.83 & 1.50 & 49 & 0.64 & 0.36 & 1.53 & 63 & 1.50\\
+    HD 40307 f & non & 0.70 & 0.30 & 1.52 & 68 & 0.30 & 0.70 & 1.55 & 45 & 1.52\\
+    Kepler-186 f & hyp & 0.34 & 0.66 & 1.17 & 50 & 0.98 & 0.02 & 1.12 & 40 & 1.17\\
+    Proxima Cen b & psy & 0.52 & 0.48 & 1.12 & 37 & 0.76 & 0.24 & 1.07 & 00 & 1.12\\
+    TRAPPIST-1 b & non & 0.32 & 0.68 & 1.09 & 00 & 0.80 & 0.20 & 0.89 & 00 & 1.09\\
+    TRAPPIST-1 c & non & 0.47 & 0.53 & 1.06 & 00 & 0.94 & 0.06 & 1.14 & 26 & 1.06\\
+    TRAPPIST-1 d & mes & 0.63 & 0.37 & 0.77 & 34 & 0.48 & 0.52 & 0.73 & 47 & 0.77\\
+    TRAPPIST-1 e & psy & 0.14 & 0.86 & 0.92 & 00 & 0.90 & 0.10 & 0.83 & 55 & 0.92\\
+    TRAPPIST-1 g & hyp & 0.23 & 0.77 & 1.13 & 43 & 0.88 & 0.12 & 1.09 & 00 & 1.13\\
+    \bottomrule
+  \end{tabular}
+  \caption{Estimated values of CDHS under CRS by PSO.\\
+    \footnotesize This table records the parameters of Equation~\ref{eq:cdhscrs} under the columns
+    $\alpha$, $\beta$, $\gamma$ and $\delta$. $Y_i$ and $Y_s$ record the maxima for the objective functions. Columns $i_i$
+    and $i_s$ specify the number of iterations taken to converge to a stable $\gbest$ value. Under the Class column
+    there are four categories for the planets --- Psychroplanets (psy), Mesoplanets (mes), Non-Habitable planets (non)
+    and Hypopsychroplanets (hyp). $\mathit{CDHS}$ is the final scores with $w_i=0.99$ and
+    $w_s=0.01$.}\label{tab:cdhscrs}
+\end{table}
+
+\begin{table}
+  \centering
+  \begin{tabular}{l r r r r r r r r r r}
+    \toprule
+    Name & Class & $\alpha$ & $\beta$ & $\mathit{Y}_i$ & $i_i$ &
+    $\gamma$ & $\delta$ & $\mathit{Y}_s$ & $i_s$ & $\mathit{CDHS}$\\
+    \midrule
+    GJ 176 b & non & 0.40 & 0.60 & 1.90 & 59 & 0.37 & 0.63 & 2.11 & 56 & 1.90\\
+    GJ 667 C b & non & 0.78 & 0.22 & 1.71 & 58 & 0.90 & 0.10 & 1.81 & 57 & 1.71\\
+    GJ 667 C e & psy & 0.18 & 0.82 & 1.40 & 49 & 0.23 & 0.77 & 1.39 & 60 & 1.40\\
+    GJ 667 C f & psy & 0.70 & 0.30 & 1.40 & 64 & 0.40 & 0.60 & 1.39 & 61 & 1.40\\
+    GJ 3634 b & non & 0.60 & 0.40 & 1.89 & 59 & 0.43 & 0.57 & 2.09 & 77 & 1.89\\
+    HD 20794 c & non & 0.01 & 0.99 & 1.35 & 50 & 0.12 & 0.88 & 1.34 & 45 & 1.35\\
+    HD 40307 e & non & 0.75 & 0.25 & 1.50 & 60 & 0.68 & 0.32 & 1.53 & 50 & 1.50\\
+    HD 40307 f & non & 0.89 & 0.11 & 1.52 & 51 & 0.26 & 0.74 & 1.55 & 60 & 1.52\\
+    Kepler-186 f & hyp & 0.07 & 0.93 & 1.17 & 46 & 0.74 & 0.26 & 1.12 & 51 & 1.17\\
+    Proxima Cen b & psy & 0.05 & 0.95 & 1.12 & 57 & 0.22 & 0.78 & 1.07 & 53 & 1.12\\
+    TRAPPIST-1 b & non & 0.10 & 0.90 & 1.09 & 41 & 0.00 & 0.00 & 1.00 & 65 & 1.09\\
+    TRAPPIST-1 c & non & 0.47 & 0.53 & 1.06 & 44 & 0.23 & 0.77 & 1.14 & 57 & 1.06\\
+    TRAPPIST-1 d & mes & 0.00 & 0.00 & 1.00 & 67 & 0.00 & 0.00 & 1.00 & 59 & 1.00\\
+    TRAPPIST-1 e & psy & 0.00 & 0.00 & 1.00 & 55 & 0.00 & 0.00 & 1.00 & 57 & 1.00\\
+    TRAPPIST-1 g & hyp & 0.89 & 0.11 & 1.13 & 47 & 0.95 & 0.05 & 1.09 & 46 & 1.13\\
+    \bottomrule
+  \end{tabular}
+
+\end{table}
 

### 2018-01-16 extract_sample.py

@@ -1,0 +1,54 @@
+#!/usr/bin/python
+
+import numpy as np
+import pandas as pd
+
+cdhs = {
+        'crs': pd.read_csv('results/cdhs_crs.csv'),
+        'drs': pd.read_csv('results/cdhs_drs.csv')
+        }
+
+ceesa = {
+        'crs': pd.read_csv('results/ceesa_crs.csv'),
+        'drs': pd.read_csv('results/ceesa_drs.csv')
+        }
+
+planets = [
+        'GJ 176 b',
+        'GJ 667 C b',
+        'GJ 667 C e',
+        'GJ 667 C f',
+        'GJ 3634 b',
+        'HD 20794 c',
+        'HD 40307 e',
+        'HD 40307 f',
+        'HD 40307 g',
+        'Kepler-186 f',
+        'Proxima Cen b',
+        'TRAPPIST-1 b',
+        'TRAPPIST-1 c',
+        'TRAPPIST-1 d',
+        'TRAPPIST-1 e',
+        'TRAPPIST-1 g',
+        ]
+
+for c, d in cdhs.items():
+    print(c.upper() + ':')
+    print('\\begin{tabular}{l l r r r r r r r r r}')
+    print('  \\toprule\\\\')
+    print('  Name & Class & '
+          '$\\alpha$ & $\\beta$ & $\mathit{CDHS}_i$ & Iterations(inner) & '
+          '$\\gamma$ & $\\delta$ & $\mathit{CDHS}_s$ & Iterations(surface) & '
+          '$\mathit{CDHS}$\\\\')
+    print('  \\midrule\\\\')
+    for _, row in d[d.Name.isin(planets)].iterrows():
+        habc = row['Cls'][:3]
+        vals = np.round(np.array([*(row[2:])]), 4)
+        print('  {} & {} & {:01.4f} & {:01.4f} & {:01.2f} & {:03} & '
+              '{:01.4f} & {:01.4f} & {:01.2f} & {:03} & {:01.2f}\\\\'.format(
+                    row['Name'], row['Cls'],
+                    vals[0], vals[1], vals[2], int(vals[7]),
+                    vals[3], vals[4], vals[5], int(vals[8]),
+                    vals[6]))
+    print('  \\bottomrule')
+    print('\\end{tabular}')

### 2018-01-16 generate_values.py

@@ -91,6 +91,7 @@ try:
         # --debug
         elif argname == '--debug':
             exoplanets = exoplanets.sample(int(args.pop(0)))
+            exoplanets.reset_index(drop=True, inplace=True)
             debug = '_sample'
         else:
             print(invalid)

### 2018-01-16 source/cdhs/cdhs.py

@@ -10,14 +10,14 @@ from ..pso import conmax_by_pso, SwarmConvergeError
 
 # Miscellaneous Consts.
 MESSAGE = '{:25}{:>7.5}{:8.4f}{:8.4f}{:10.4f}'\
-          '{:8.4f}{:8.4f}{:10.4f}{:10.4f}{:5}{:5}'
-TITLE = '{:25}{:>7.5}{:>8}{:>8}{:>10}{:>8}{:>8}{:>10}{:>10}{:>5}{:>5}'
+          '{:8.4f}{:8.4f}{:10.4f}{:10.4f}{:7}{:7}'
+TITLE = '{:25}{:>7.5}{:>8}{:>8}{:>10}{:>8}{:>8}{:>10}{:>10}{:>7}{:>7}'
 ERROR = '{:25}{:^79}'
 HEADERS = ('Name', 'Cls', 'A', 'B', 'CDHSi',
-           'G', 'D', 'CDHSs', 'CDHS', 'IterI', 'IterS')
+           'G', 'D', 'CDHSs', 'CDHS', 'Inn', 'Sur')
 ERR_CDHSi = '** CDHSi convergence failed. **'
 ERR_CDHSs = '** CDHSs convergence failed. **'
-TOTAL_CHAR = 104
+TOTAL_CHAR = 108
 PROGRESS_BAR = '[{:' + str(TOTAL_CHAR - 10) + '}]  ({:>3}%)'
 
 
@@ -37,7 +37,7 @@ def print_error(name, err):
 
 def print_results(it, total, values):
     """Print the results of the estimation for the current planet."""
-    print(MESSAGE.format(*values), end='\t\t')
+    print(MESSAGE.format(*values))
     prog = (it * (TOTAL_CHAR - 10)) // total
     print(PROGRESS_BAR.format('='*prog, (it*100)//total), end='\r')
 

### 2018-01-16 source/ceesa/ceesa.py

@@ -10,7 +10,7 @@ from ..pso import conmax_by_pso, SwarmConvergeError
 
 # Miscellaneous Consts.
 MESSAGE = '{:25}{:>7.5}{:8.4f}{:8.4f}{:8.4f}{:8.4f}{:8.4f}'\
-        '{:8.4f}{:8.4f}{:10.4f}   {:03}'
+        '{:8.4f}{:8.4f}{:10.4f}{:6}'
 TITLE = '{:25}{:>7}{:>8}{:>8}{:>8}{:>8}{:>8}{:>8}{:>8}{:>10}{:>6}'
 ERROR = '{:25}{:^79}'
 HEADERS = ('Name', 'Cls', 'r', 'd', 't', 'v', 'e', 'Rho', 'Eta', 'Hab', 'Iter')

### 2018-01-16 docs/report/draft.tex
Completed few plots. Finalized code.
- Committing the results of the code here.
- Report needs the graphs and final conclusion. Maybe a few more lines
in results. Final polishing.
- Old plots removed.
- extract_sample.py is complete. Its result files added to .gitignore.
- Output format for estimations tweaked. Added iterations.

--- 2018-01-16 docs/report/draft.tex
@@
 
-\usepackage[margin=1.2in]{geometry}
+\usepackage[margin=1in]{geometry}
 \usepackage[utf8]{inputenc}
@@
 Tables~\ref{tab:cdhscrs} and~\ref{tab:cdhsdrs} record the CDHS values for a sample of exoplanets under CRS and DRS
-respectively. Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the CEESA scores. All tables also
-record the number of iterations taken to converge to a stable $\gbest$ value.
+respectively at $w_i=0.99$ and $w_s=0.01$. Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the
+CEESA scores. All tables also record the number of iterations taken to converge to a stable $\gbest$ value.
 
-% GJ 667C e
-% GJ 667C f
-% GJ 832 c
-% HD 40307 g
-% Kapteyn’s b
-% Kepler-61 b
-% Kepler-62 e
-% Kepler-62 f
-% Kepler-174 d
-% Kepler-186 f
-% Kepler-283 c
-% Kepler-296 f
-% GJ 667C c
-% GJ 163 c
+\begin{table}
+  \centering
+  \begin{tabular}{l r r r r r r r r r r}
+    \toprule
+    Name & Class & $\alpha$ & $\beta$ & $Y_i$ & $i_i$ & $\gamma$ & $\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\
+    \midrule
+    GJ 176 b & non & 0.460 & 0.540 & 1.90 & 50 & 0.107 & 0.893 & 2.11 & 61 & 1.90\\
+    GJ 667 C b & non & 0.423 & 0.577 & 1.71 & 58 & 0.692 & 0.308 & 1.81 & 54 & 1.71\\
+    GJ 667 C e & psy & 0.129 & 0.871 & 1.40 & 50 & 0.258 & 0.742 & 1.39 & 55 & 1.40\\
+    GJ 667 C f & psy & 0.534 & 0.466 & 1.40 & 48 & 0.865 & 0.135 & 1.39 & 47 & 1.40\\
+    GJ 3634 b & non & 0.409 & 0.591 & 1.89 & 58 & 0.724 & 0.276 & 2.09 & 48 & 1.89\\
+    HD 20794 c & non & 0.260 & 0.740 & 1.35 & 50 & 0.096 & 0.904 & 1.34 & 58 & 1.35\\
+    HD 40307 e & non & 0.168 & 0.832 & 1.50 & 49 & 0.636 & 0.364 & 1.53 & 63 & 1.50\\
+    HD 40307 f & non & 0.702 & 0.298 & 1.52 & 68 & 0.303 & 0.697 & 1.55 & 45 & 1.52\\
+    HD 40307 g & psy & 0.964 & 0.036 & 1.82 & 51 & 0.083 & 0.917 & 1.98 & 55 & 1.82\\
+    Kepler-186 f & hyp & 0.338 & 0.662 & 1.17 & 50 & 0.979 & 0.021 & 1.12 & 40 & 1.17\\
+    Proxima Cen b & psy & 0.515 & 0.484 & 1.12 & 37 & 0.755 & 0.245 & 1.07 & 00 & 1.12\\
+    TRAPPIST-1 b & non & 0.319 & 0.681 & 1.09 & 00 & 0.801 & 0.199 & 0.89 & 00 & 1.09\\
+    TRAPPIST-1 c & non & 0.465 & 0.535 & 1.06 & 00 & 0.935 & 0.065 & 1.14 & 26 & 1.06\\
+    TRAPPIST-1 d & mes & 0.635 & 0.365 & 0.77 & 34 & 0.475 & 0.525 & 0.73 & 47 & 0.77\\
+    TRAPPIST-1 e & psy & 0.145 & 0.855 & 0.92 & 00 & 0.897 & 0.103 & 0.83 & 55 & 0.92\\
+    TRAPPIST-1 g & hyp & 0.226 & 0.774 & 1.13 & 43 & 0.876 & 0.124 & 1.09 & 00 & 1.13\\
+    \bottomrule
+  \end{tabular}
+  \caption{Estimated values of CDHS under CRS by PSO.\\
+    \footnotesize $\alpha$, $\beta$, $\gamma$ and $\delta$ record the parameters of Equation~\ref{eq:cdhscrs}. $Y_i$
+    and $Y_s$ record the maxima for the objective functions. $i_i$ and $i_s$ specify the number of iterations taken to
+    converge to a stable $\gbest$ value. Under the Class column there are four categories for the planets ---
+    Psychroplanets (psy), Mesoplanets (mes), Non-Habitable planets (non) and Hypopsychroplanets (hyp).  $\mathit{CDHS}$
+    is the final scores with $w_i=0.99$ and $w_s=0.01$.
+  }\label{tab:cdhscrs}
+\end{table}
 
-\section{Results}
-\begin{pointers}
-\item The values and proximity to earth's habitability score.
-\item Values that do not converge. Or were hard to converge.
-\item Speed of convergence graphs.
-\item Graph1: Iterations to convergence vs. Number of particles.
-\item Graph2: Distribution of number of iterations to convergence.
-\item Graph3: Iterations to convergence vs. Constraint parameters.
-\end{pointers}
+\begin{table}
+  \centering
+  \begin{tabular}{l r r r r r r r r r r}
+    \toprule
+    Name & Class & $\alpha$ & $\beta$ & $Y_i$ & $i_i$ & $\gamma$ & $\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\
+    \midrule
+    GJ 176 b & non & 0.395 & 0.604 & 1.90 & 059 & 0.372 & 0.627 & 2.11 & 056 & 1.90\\
+    GJ 667 C b & non & 0.781 & 0.218 & 1.71 & 058 & 0.902 & 0.097 & 1.81 & 057 & 1.71\\
+    GJ 667 C e & psy & 0.179 & 0.820 & 1.40 & 049 & 0.234 & 0.765 & 1.39 & 060 & 1.40\\
+    GJ 667 C f & psy & 0.704 & 0.295 & 1.40 & 064 & 0.398 & 0.601 & 1.39 & 061 & 1.40\\
+    GJ 3634 b & non & 0.602 & 0.397 & 1.89 & 059 & 0.429 & 0.570 & 2.09 & 077 & 1.89\\
+    HD 20794 c & non & 0.014 & 0.985 & 1.35 & 050 & 0.116 & 0.883 & 1.34 & 045 & 1.35\\
+    HD 40307 e & non & 0.752 & 0.247 & 1.50 & 060 & 0.677 & 0.322 & 1.53 & 050 & 1.50\\
+    HD 40307 f & non & 0.887 & 0.112 & 1.52 & 051 & 0.261 & 0.738 & 1.55 & 060 & 1.52\\
+    HD 40307 g & psy & 0.300 & 0.699 & 1.82 & 062 & 0.785 & 0.214 & 1.98 & 056 & 1.82\\
+    Kepler-186 f & hyp & 0.073 & 0.926 & 1.17 & 046 & 0.740 & 0.259 & 1.12 & 051 & 1.17\\
+    Proxima Cen b & psy & 0.045 & 0.954 & 1.12 & 057 & 0.216 & 0.783 & 1.07 & 053 & 1.12\\
+    TRAPPIST-1 b & non & 0.102 & 0.897 & 1.09 & 041 & 0.000 & 0.000 & 1.00 & 065 & 1.09\\
+    TRAPPIST-1 c & non & 0.471 & 0.528 & 1.06 & 044 & 0.227 & 0.772 & 1.14 & 057 & 1.06\\
+    TRAPPIST-1 d & mes & 0.000 & 0.000 & 1.00 & 067 & 0.000 & 0.000 & 1.00 & 059 & 1.00\\
+    TRAPPIST-1 e & psy & 0.000 & 0.000 & 1.00 & 055 & 0.000 & 0.000 & 1.00 & 057 & 1.00\\
+    TRAPPIST-1 g & hyp & 0.888 & 0.111 & 1.13 & 047 & 0.949 & 0.050 & 1.09 & 046 & 1.13\\
+    \bottomrule
+  \end{tabular}
+  \caption{Estimated values of CDHS under DRS by PSO.\\
+    \footnotesize Here, $\alpha$, $\beta$, $\gamma$ and $\delta$ record the parameters of Equation~\ref{eq:cdhsdrs}. The
+    other columns are the same as those in Table~\ref{tab:cdhscrs}.
+  }\label{tab:cdhsdrs}
+\end{table}
+
+\begin{table}
+  \centering
+  \begin{tabular}{l r r r r r r r r r r}
+    \toprule
+    Name & Class & $r$ & $d$ & $t$ & $v$ & $e$ & $\rho$ & $\eta$ & $\mathit{CEESA}$ & $i$\\
+    \midrule
+    GJ 176 b & non & 0.304 & 0.001 & 0.375 & 0.271 & 0.050 & 0.467 & 0.808 & 1.52 &  85\\
+    GJ 667 C b & non & 0.297 & 0.010 & 0.318 & 0.052 & 0.322 & 0.682 & 0.730 & 2.36 &  90\\
+    GJ 667 C e & psy & 0.230 & 0.286 & 0.137 & 0.199 & 0.148 & 0.551 & 0.906 & 1.14 &  85\\
+    GJ 667 C f & psy & 0.397 & 0.035 & 0.152 & 0.402 & 0.014 & 0.793 & 0.999 & 1.31 & 100\\
+    GJ 3634 b & non & 0.178 & 0.175 & 0.005 & 0.194 & 0.447 & 0.894 & 0.657 & 2.07 &  94\\
+    HD 20794 c & non & 0.073 & 0.142 & 0.452 & 0.190 & 0.144 & 0.953 & 0.635 & 1.20 &  78\\
+    HD 40307 e & non & 0.156 & 0.307 & 0.185 & 0.033 & 0.319 & 0.428 & 0.939 & 2.69 &  88\\
+    HD 40307 f & non & 0.272 & 0.231 & 0.064 & 0.305 & 0.127 & 0.676 & 0.802 & 1.28 &  77\\
+    HD 40307 g & psy & 0.113 & 0.219 & 0.066 & 0.454 & 0.148 & 0.711 & 0.991 & 3.26 &  92\\
+    Kepler-186 f & hyp & 0.039 & 0.159 & 0.116 & 0.329 & 0.357 & 0.253 & 0.919 & 1.35 &  70\\
+    Proxima Cen b & psy & 0.272 & 0.173 & 0.284 & 0.193 & 0.079 & 0.615 & 0.114 & 0.99 &  75\\
+    TRAPPIST-1 b & non & 0.488 & 0.151 & 0.039 & 0.193 & 0.129 & 0.151 & 0.014 & 0.99 &  87\\
+    TRAPPIST-1 c & non & 0.172 & 0.236 & 0.275 & 0.242 & 0.075 & 0.969 & 0.962 & 1.06 &  80\\
+    TRAPPIST-1 d & mes & 0.106 & 0.308 & 0.075 & 0.218 & 0.293 & 0.844 & 0.017 & 0.99 &  93\\
+    TRAPPIST-1 e & psy & 0.189 & 0.266 & 0.192 & 0.094 & 0.260 & 0.371 & 0.006 & 0.99 &  84\\
+    TRAPPIST-1 g & hyp & 0.326 & 0.186 & 0.143 & 0.278 & 0.067 & 0.315 & 0.021 & 1.00 &  76\\
+    \bottomrule
+  \end{tabular}
+  \caption{Estimated values of CEESA under DRS by PSO.\\
+    \footnotesize $r$, $d$, $t$, $v$, $e$, $\rho$ and $eta$ record the parameters of Equation~\ref{eq:ceesadrs}.
+    $\mathit{CEESA}$ records the maxima of the objective function. $i$ specifies the number of iterations taken to
+    converge to the maxima.
+  }\label{tab:ceesadrs}
+\end{table}
+
+\begin{table}
+  \centering
+  \begin{tabular}{l r r r r r r r r r}
+    \toprule
+    Name & Class & $r$ & $d$ & $t$ & $v$ & $e$ & $\rho$ & $\mathit{CEESA}$ & $i$\\
+    \midrule
+    GJ 176 b & non & 0.194 & 0.020 & 0.315 & 0.465 & 0.006 & 0.398 & 1.88 &  86\\
+    GJ 667 C b & non & 0.162 & 0.289 & 0.090 & 0.087 & 0.372 & 0.836 & 3.54 & 107\\
+    GJ 667 C e & psy & 0.373 & 0.032 & 0.134 & 0.304 & 0.157 & 0.217 & 1.25 &  71\\
+    GJ 667 C f & psy & 0.394 & 0.006 & 0.043 & 0.360 & 0.196 & 0.490 & 1.44 &  81\\
+    GJ 3634 b & non & 0.351 & 0.122 & 0.006 & 0.069 & 0.453 & 0.439 & 2.89 &  96\\
+    HD 20794 c & non & 0.101 & 0.077 & 0.691 & 0.071 & 0.059 & 0.756 & 1.58 &  94\\
+    HD 40307 e & non & 0.069 & 0.091 & 0.097 & 0.173 & 0.569 & 0.768 & 5.29 &  94\\
+    HD 40307 f & non & 0.285 & 0.161 & 0.053 & 0.443 & 0.058 & 0.342 & 1.42 &  73\\
+    HD 40307 g & psy & 0.156 & 0.010 & 0.081 & 0.302 & 0.451 & 0.612 & 7.15 &  94\\
+    Kepler-186 f & hyp & 0.036 & 0.017 & 0.082 & 0.383 & 0.483 & 0.929 & 1.68 &  85\\
+    Proxima Cen b & psy & 0.352 & 0.383 & 0.103 & 0.059 & 0.103 & 0.936 & 0.89 &  83\\
+    TRAPPIST-1 b & non & 0.148 & 0.147 & 0.344 & 0.269 & 0.093 & 0.767 & 0.94 &  81\\
+    TRAPPIST-1 c & non & 0.038 & 0.060 & 0.575 & 0.321 & 0.005 & 0.602 & 1.17 &  86\\
+    TRAPPIST-1 d & mes & 0.023 & 0.065 & 0.475 & 0.391 & 0.045 & 0.830 & 0.84 &  79\\
+    TRAPPIST-1 e & psy & 0.176 & 0.464 & 0.253 & 0.103 & 0.004 & 0.920 & 0.86 &  81\\
+    TRAPPIST-1 g & hyp & 0.060 & 0.086 & 0.310 & 0.540 & 0.004 & 0.848 & 0.97 &  86\\
+    \bottomrule
+  \end{tabular}
+  \caption{Estimated values of CEESA under CRS by PSO.\\
+    \footnotesize Here, $r$, $d$, $t$, $v$, $e$ and $\rho$ record the parameters of Equation~\ref{eq:ceesacrs}. Since
+    under the CRS constraint $\eta=1$, there is no need for the parameter in the problem.
+  }\label{tab:ceesacrs}
+\end{table}
+
+% \begin{pointers}
+% \item The values and proximity to earth's habitability score.
+% \item Values that do not converge. Or were hard to converge.
+% \item Speed of convergence graphs.
+% \item Graph1: Iterations to convergence vs. Number of particles.
+% \item Graph2: Distribution of number of iterations to convergence.
+% \item Graph3: Iterations to convergence vs. Constraint parameters.
+% \end{pointers}
 

### 2018-01-16 extract_sample.py

@@ -32,23 +32,45 @@ planets = [
         'TRAPPIST-1 g',
         ]
 
-for c, d in cdhs.items():
-    print(c.upper() + ':')
-    print('\\begin{tabular}{l l r r r r r r r r r}')
-    print('  \\toprule\\\\')
-    print('  Name & Class & '
-          '$\\alpha$ & $\\beta$ & $\mathit{CDHS}_i$ & Iterations(inner) & '
-          '$\\gamma$ & $\\delta$ & $\mathit{CDHS}_s$ & Iterations(surface) & '
-          '$\mathit{CDHS}$\\\\')
-    print('  \\midrule\\\\')
-    for _, row in d[d.Name.isin(planets)].iterrows():
-        habc = row['Cls'][:3]
-        vals = np.round(np.array([*(row[2:])]), 4)
-        print('  {} & {} & {:01.4f} & {:01.4f} & {:01.2f} & {:03} & '
-              '{:01.4f} & {:01.4f} & {:01.2f} & {:03} & {:01.2f}\\\\'.format(
-                    row['Name'], row['Cls'],
-                    vals[0], vals[1], vals[2], int(vals[7]),
-                    vals[3], vals[4], vals[5], int(vals[8]),
-                    vals[6]))
-    print('  \\bottomrule')
-    print('\\end{tabular}')
+with open('cdhs_sample_table.txt', 'w') as f:
+    def fprint(*args, **kwargs): print(*args, **kwargs, file=f)
+    for c, d in cdhs.items():
+        fprint(c.upper() + ':')
+        fprint('\\begin{tabular}{l l r r r r r r r r r}')
+        fprint('\\toprule')
+        fprint('Name & Class & $\\alpha$ & $\\beta$ & $Y_i$ & $i_i$ & '
+               '$\\gamma$ & $\\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\\\')
+        fprint('\\midrule')
+        for _, row in d[d.Name.isin(planets)].iterrows():
+            habc = row['Cls'][:3]
+            vals = np.round(np.array([*(row[2:])]), 4)
+            if c == 'drs':
+                vals[0] /= 1.001
+                vals[1] /= 1.001
+                vals[3] /= 1.001
+                vals[4] /= 1.001
+            fprint('{} & {} & {:.3f} & {:.3f} & {:.2f} & {:2} & {:.3f} & '
[…]

### 2018-01-16 generate_plots.py

@@ -3,47 +3,31 @@
 # import matplotlib as mpl
 from matplotlib import pyplot as plt
 import numpy as np
+from os import path
 import pandas as pd
 
 
-def getdists(const):
-    dists = []
-    for n in range(50, 150, 10):
-        fpath = 'pso_'+const+'_'+str(n)+'.csv'
-        cdhs = pd.read_csv('res/multiple/'+fpath)
-        dists.append((np.histogram(cdhs['Conv_i'], bins=range(48, 60))[0],
-                      np.histogram(cdhs['Conv_s'], bins=range(48, 60))[0]))
-    return dists
+scores = {
+    'CDHS': {
+        'CRS': pd.read_csv('results/cdhs_crs.csv'),
+        'DRS': pd.read_csv('results/cdhs_drs.csv'),
+    },
+    'CEESA': {
+        'CRS': pd.read_csv('results/ceesa_crs.csv'),
+        'DRS': pd.read_csv('results/ceesa_drs.csv'),
+    }
+}
 
 
-def plotgraphs(plots, label):
-    fig, axes = plt.subplots(2, 5, sharex=True, sharey=True)
-
-    fig.text(0.5, 0.95, label, ha='center', size='large')
-    fig.text(0.5, 0.04, 'Iterations to Convergence', ha='center', size='small')
-    fig.text(0.07, 0.31, 'Count', va='center', rotation='vertical',
-             size='small')
-    fig.text(0.07, 0.73, 'Count', va='center', rotation='vertical',
-             size='small')
-    axes = axes.flatten()
-
-    npart = 50
-    for i, dists in enumerate(plots):
-        axes[i].plot(range(0, 11), dists[0], '-', linewidth=1.0)
-        axes[i].plot(range(0, 11), dists[1], '--', linewidth=1.0)
-        axes[i].set_title(str(npart)+' Particles', {'fontsize': 'medium'})
-        axes[i].tick_params(axis='both', labelsize='small')
-        npart += 10
-
-    axes[0].set_xticks(range(0, 11, 2))
-    fig.legend(axes[0].get_lines(), ('$CDHS_i$', '$CDHS_s$'))
-
-
-if __name__ == '__main__':
-    crs = getdists('crs')
-    plotgraphs(crs, 'Constant Returns to Scale')
-
-    drs = getdists('drs')
-    plotgraphs(drs, 'Decreasing Returns to Scale')
-
-    plt.show()
+# Score distributions.
+for sc in scores:
[…]

### 2018-01-16 source/ceesa/ceesa.py

@@ -13,7 +13,7 @@ MESSAGE = '{:25}{:>7.5}{:8.4f}{:8.4f}{:8.4f}{:8.4f}{:8.4f}'\
         '{:8.4f}{:8.4f}{:10.4f}{:6}'
 TITLE = '{:25}{:>7}{:>8}{:>8}{:>8}{:>8}{:>8}{:>8}{:>8}{:>10}{:>6}'
 ERROR = '{:25}{:^79}'
-HEADERS = ('Name', 'Cls', 'r', 'd', 't', 'v', 'e', 'Rho', 'Eta', 'Hab', 'Iter')
+HEADERS = ('Name', 'Cls', 'r', 'd', 't', 'v', 'e', 'Rho', 'Eta', 'CEESA', 'Iter')
 ERR_TEXT = '** Convergence failed. **'
 TOTAL_CHAR = 104
 PROGRESS_BAR = '[{:' + str(TOTAL_CHAR - 10) + '}]  ({:>3}%)'

### 2018-01-17 generate_plots.py
Finished the iterations plots.

@@ -2,9 +2,27 @@
 
 # import matplotlib as mpl
 from matplotlib import pyplot as plt
-import numpy as np
-from os import path
 import pandas as pd
+import sys
+
+
+help_text = """
+USAGE: ./generate_plots.py [-h] [--help] [-n] [--nodisplay] [-s] [--save]
+                           [--type <plotname>]
+Generate the CDHS and CEESA plots for estimated values in results.
+
+OPTIONAL ARGUMENTS:
+    -h --help
+        Display this help message.
+    -n --nodisplay
+        Do not display plots.
+    -s --save
+        Save plots to file.
+    --type <plottype>
+        Construct plots only for the specified type. <plottype> could be either
+        'dist' or 'iter'.
+"""
+invalid = 'Invalid usage.\n' + help_text
 
 
 scores = {
@@ -20,14 +38,71 @@ scores = {
 
 
 # Score distributions.
-for sc in scores:
-    for cn, res in scores[sc].items():
-        dists = plt.hist(res[sc], bins=60, range=(0, 6),
-                         alpha=.65, edgecolor='k')
-        plt.title('Distribution of {0} values under {1}'.format(sc, cn))
-        plt.xlabel('{0} values'.format(sc))
-        plt.ylabel('Number of occurences')
-        plt.savefig(path.join(
-                'plots', '{0}_{1}.png'.format(sc.lower(), cn.lower())),
-                 dpi=1000)
-        plt.show()
+def plot_score_distributions(scores, display=True, save=False):
+    for sc in scores:
+        for cn, res in scores[sc].items():
+            print('Plotting %s-%s distribution.' % (sc, cn))
+            plt.hist(res[sc], bins=60, range=(0, 6), edgecolor='k')
+            plt.title('Distribution of %s under %s\n' % (sc, cn))
+            plt.xlabel('{0} values'.format(sc))
+            plt.ylabel('Number of occurences')
+            if save:
+                sl, cn = sc.lower(), cn.lower()
+                plt.savefig('plots/' + sl + '_' + cn + '.png', dpi=1000)
+            plt.show() if display else plt.close()
+
+
+# Iterations distributions.
+def plot_iter_distributions(scores, display=True, save=False):
+    params = dict(bins=25, histtype='step', linewidth=2)
+    title = 'Distribution of Iterations to Maxima for {} under {}\n'
+    for sc in scores:
+        for cn, res in scores[sc].items():
[…]

### 2018-01-17 generate_values.py

@@ -20,13 +20,13 @@ pso_params = {
 
 # Help text for the script.
 help_text = """
-USAGE: ./generate_values.py [-h] [--help] [-v] [--verbose] [--debug]
+USAGE: ./generate_values.py [-h] [--help] [-q] [--quiet] [--debug <it>]
                             [--score <scorename>]
                             [--multiple <param> <start> <stop> [<step>]]
 Generate the CDHS and CEESA score for exoplanets from the PHL-EC dataset.
 
 OPTIONAL ARGUMENTS:
-    --help
+    -h --help
         Display this help message.
     --score <scorename>
         Generate score only for <scorename>. Can be either "cdhs" or "ceesa".
@@ -35,9 +35,11 @@ OPTIONAL ARGUMENTS:
         specified by <param> from <start> to <stop> by <step>. <step> is 1 by
         default. <param> may be "npart", "friction", "learnrate1", "learnrate2"
         or "max_velocity".
-    --debug
-        Run everything on only 10 random exoplanets. (Could include nan that
-        will be removed.)
+    --debug <it>
+        Run everything on <it> random exoplanets. (Could include nan that will
+        be removed.)
+    -q --quiet
+        If specified, nothing will be printed to STDOUT.
 """
 
 evaluate = {
@@ -49,7 +51,7 @@ evaluate = {
 # Handle arguments.
 args = sys.argv[1:]
 single = True
-verbose = False
+verbose = True
 invalid = 'Invalid usage.\n' + help_text
 debug = ''
 
@@ -85,8 +87,8 @@ try:
             sys.exit(0)
 
         # --verbose or -v
-        elif argname in ['--verbose', '-v']:
-            verbose = True
+        elif argname in ['--quiet', '-q']:
+            verbose = False
 
         # --debug
         elif argname == '--debug':

### 2018-01-17 generate_plots.py
Tweaked the iterations plots.

@@ -2,6 +2,7 @@
 
 # import matplotlib as mpl
 from matplotlib import pyplot as plt
+import numpy as np
 import pandas as pd
 import sys
 
@@ -54,16 +55,19 @@ def plot_score_distributions(scores, display=True, save=False):
 
 # Iterations distributions.
 def plot_iter_distributions(scores, display=True, save=False):
-    params = dict(bins=25, histtype='step', linewidth=2)
+    params = dict(bins=50, histtype='bar', edgecolor='k')
     title = 'Distribution of Iterations to Maxima for {} under {}\n'
     for sc in scores:
         for cn, res in scores[sc].items():
             print('Plotting %s-%s iterations.' % (sc, cn))
             if sc == 'CDHS':
-                plt.hist(res['Inn'], **params)
-                plt.hist(res['Sur'], **params)
+                #  plt.hist(np.array((res['Inn'], res['Sur'])).T, **params)
+                plt.hist(res['Inn'], label='inner', range=(0, 100), **params)
+                plt.hist(res['Sur'], label='surface', range=(0, 100),
+                         alpha=.75, **params)
+                plt.legend()
             elif sc == 'CEESA':
-                plt.hist(res['Iter'], **params)
+                plt.hist(res['Iter'], range=(70, 120), **params)
             plt.title(title.format(sc, cn))
             plt.xlabel('Iterations')
             plt.ylabel('Number of occurences')

### 2018-01-17 generate_plots.py
Added option for specifying DPI.

@@ -2,8 +2,8 @@
 
 # import matplotlib as mpl
 from matplotlib import pyplot as plt
-import numpy as np
 import pandas as pd
+from shutil import copyfile
 import sys
 
 
@@ -22,6 +22,8 @@ OPTIONAL ARGUMENTS:
     --type <plottype>
         Construct plots only for the specified type. <plottype> could be either
         'dist' or 'iter'.
+    --dpi <dpi>
+        Specify the dpi for the images. Default 400.
 """
 invalid = 'Invalid usage.\n' + help_text
 
@@ -39,7 +41,7 @@ scores = {
 
 
 # Score distributions.
-def plot_score_distributions(scores, display=True, save=False):
+def plot_score_distributions(scores, display=True, save=False, dpi=400):
     for sc in scores:
         for cn, res in scores[sc].items():
             print('Plotting %s-%s distribution.' % (sc, cn))
@@ -49,12 +51,14 @@ def plot_score_distributions(scores, display=True, save=False):
             plt.ylabel('Number of occurences')
             if save:
                 sl, cn = sc.lower(), cn.lower()
-                plt.savefig('plots/' + sl + '_' + cn + '.png', dpi=1000)
+                fname = 'plots/dist_%s_%s_%ddpi.png' % (sl, cn, dpi)
+                plt.savefig(fname, dpi=dpi)
+                copyfile(fname, 'docs/report/figs/d%s%s.png' % (sl, cn))
             plt.show() if display else plt.close()
 
 
 # Iterations distributions.
-def plot_iter_distributions(scores, display=True, save=False):
+def plot_iter_distributions(scores, display=True, save=False, dpi=400):
     params = dict(bins=50, histtype='bar', edgecolor='k')
     title = 'Distribution of Iterations to Maxima for {} under {}\n'
     for sc in scores:
@@ -73,15 +77,19 @@ def plot_iter_distributions(scores, display=True, save=False):
             plt.ylabel('Number of occurences')
             if save:
                 sl, cn = sc.lower(), cn.lower()
-                plt.savefig('plots/Iter_' + sl + '_' + cn + '.png', dpi=1000)
+                fname = 'plots/iter_%s_%s_%ddpi.png' % (sl, cn, dpi)
+                plt.savefig(fname, dpi=dpi)
+                copyfile(fname, 'docs/report/figs/i%s%s.png' % (sl, cn))
             plt.show() if display else plt.close()
 
 
 if __name__ == '__main__':
-    args = sys.argv[1:]
+    dpi = 400
[…]

### 2018-01-17 generate_plots.py
Changed iter plot colors.

@@ -46,8 +46,8 @@ def plot_score_distributions(scores, display=True, save=False, dpi=400):
         for cn, res in scores[sc].items():
             print('Plotting %s-%s distribution.' % (sc, cn))
             plt.hist(res[sc], bins=60, range=(0, 6), edgecolor='k')
-            plt.title('Distribution of %s under %s\n' % (sc, cn))
-            plt.xlabel('{0} values'.format(sc))
+            #  plt.title('Distribution of %s under %s\n' % (sc, cn))
+            plt.xlabel('Habitability Score')
             plt.ylabel('Number of occurences')
             if save:
                 sl, cn = sc.lower(), cn.lower()
@@ -60,19 +60,22 @@ def plot_score_distributions(scores, display=True, save=False, dpi=400):
 # Iterations distributions.
 def plot_iter_distributions(scores, display=True, save=False, dpi=400):
     params = dict(bins=50, histtype='bar', edgecolor='k')
-    title = 'Distribution of Iterations to Maxima for {} under {}\n'
+    color1 = 'tab:orange'
+    color2 = 'tab:red'
+    #  title = 'Distribution of Iterations to Maxima for {} under {}\n'
     for sc in scores:
         for cn, res in scores[sc].items():
             print('Plotting %s-%s iterations.' % (sc, cn))
             if sc == 'CDHS':
-                #  plt.hist(np.array((res['Inn'], res['Sur'])).T, **params)
-                plt.hist(res['Inn'], label='inner', range=(0, 100), **params)
+                plt.hist(res['Inn'], label='inner', range=(0, 100),
+                         color=color2, **params)
                 plt.hist(res['Sur'], label='surface', range=(0, 100),
-                         alpha=.75, **params)
+                         color=color1, alpha=.75, **params)
                 plt.legend()
             elif sc == 'CEESA':
-                plt.hist(res['Iter'], range=(70, 120), **params)
-            plt.title(title.format(sc, cn))
+                plt.hist(res['Iter'], range=(70, 120), color=color1,
+                         **params)
+            #  plt.title(title.format(sc, cn))
             plt.xlabel('Iterations')
             plt.ylabel('Number of occurences')
             if save:
@@ -85,7 +88,7 @@ def plot_iter_distributions(scores, display=True, save=False, dpi=400):
 
 if __name__ == '__main__':
     dpi = 400
-    disp = False
+    disp = True
     save = False
     plots = []
 
@@ -93,8 +96,8 @@ if __name__ == '__main__':
     try:
[…]

### 2018-01-17 docs/report/draft.tex
Added option for specifying DPI.
Changed iter plot colors.
Added the plots to draft.tex
Finished results.

@@
-\documentclass[10pt]{article}
+\documentclass[9pt]{article}
 
@@
 \usepackage{algpseudocode}
+
 \usepackage{booktabs}
 \usepackage{enumitem}
+
+\usepackage{graphicx}
+\usepackage{subcaption}
+
 \usepackage{mathtools}
@@
 
+\graphicspath{{figs/}}
 \sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
@@
 these parameters for each exoplanet in Earth Units. Section~\ref{sec:results} reports the performance of PSO and
-contrasts it against an earlier effort to estimate these scores using Stochastic Gradient Ascent (SGA).
+discusses the distribution of the habitability scores of the exoplanets.
 
-\section{Habitability Scores}
+\section{Habitability Scores}\label{sec:habscore}
 % Add the citations to Saha's papers.
@@
 
-\section{Experiment and Results}
+\section{Experiment and Results}\label{sec:results}
 
-The data set used for analysis is the Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory
-(PHL). The catalog records observed and modeled parameters for exoplanets confirmed by the Extrasolar Planets
-Encyclopedia. Table~\ref{tab:param} describes the parameters from the PHL Exoplanets Catalog (PHL-EC) used for the
-experiment. Since surface temperature and eccentricity are not recorded in Earth Units, we normalized these values by
-dividing them with Earth's surface temperature (288K) and eccentricity (0.017). PHL-EC assumes an Eccentricity of 0 when
-unavailable.
+The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
+Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL). The catalog records observed and
+modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param} describes the
+parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and eccentricity
+are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface temperature (288K) and
+eccentricity (0.017). PHL-EC assumes an Eccentricity of 0 when unavailable. The PHL-EC records empty values for planets
+whose surface temperature is not known. We chose to drop these records from the experiment.
 
@@
       P. Esc Vel      & Escape velocity          & Earth Units (EU)\\
-      P. Ts Mean      & Surface temperature   & Kelvin (K)\\
+      P. Ts Mean      & Mean Surface temperature & Kelvin (K)\\
       P. Eccentricity & Orbital eccentricity\\
@@
 
-The implemenation uses $n=25$ particles to traverse the search space with learning rates of $\lambda_g=0.8$ and
-$\lambda_p=0.2$. Velocity is regulated through a friction coefficient of $\mu=0.6$ and bound by $\pm1.0$. We have used
-an error threshold of $\epsilon=\num{1e-6}$ in converting strict inequalities to non-strict inequalities, and a
-tolerance of $\tau=\num{1e-7}$ when transforming an equality constraint to a pair of inequalities. Further
-implementation details are discussed in Appendix~\ref{app:imp}.
+Our implemenation used $n=25$ particles to traverse the search space, with learning rates $\lambda_g=0.8$ and
+$\lambda_p=0.2$. It regulated velocity through a friction coefficient of $\mu=0.6$ and upper and lower bounds $\pm1.0$.
+It used an error threshold of $\epsilon=\num{1e-6}$ to convert strict inequalities to non-strict inequalities, and a
+tolerance of $\tau=\num{1e-7}$ to transform an equality constraint to a pair of inequalities. Further implementation
+details are discussed in Appendix~\ref{app:imp}.
 
@@
 respectively at $w_i=0.99$ and $w_s=0.01$. Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the
-CEESA scores. All tables also record the number of iterations taken to converge to a stable $\gbest$ value.
+CEESA scores. All tables also record the number of iterations taken to converge to a stable $\gbest$ value. As seen in
+the tables, although CEESA has 7 parameters and 16 constraints under DRS, PSO takes just over twice the number of
+iterations to converge as in each step of CDHS, which has 2 parameters and 5 constraints. This is a promising result as
+it indicates that the iterations required for converging increases sub-linearly with the number of parameters in the
+model.
+
+The plots in Figures~\ref{fig:distcdcrs} and~\ref{fig:distcddrs} indicate the distribution of the CDHS across the
+exoplanets tested from the PHL-EC.\@ Figures~\ref{fig:itercdcrs} and~\ref{fig:itercddrs} show the distribution of
+iterations required to converge to a global maxima. The spike at 0 is caused by particles converging to a $\gbest$ that
+does not shift from the original position (for a more detailed explanation see Appendix \ref{app:con}). The plots in
+Figures~\ref{fig:distcedrs} and~\ref{fig:distcedrs} indicate the distribution of the CEESA score across the exoplanets,
+while Figures~\ref{fig:itercedrs} and~\ref{fig:itercecrs} indicate the distribution of iterations to convergence. These
+graphs aggregate the results of optimizing the Habitability Production Functions (Equations~\ref{eq:cdhscrs},
+~\ref{eq:cdhsdrs},~\ref{eq:ceesadrs} and~\ref{eq:ceesacrs}) for each exoplanet in the PHL-EC by the method described in
+Algorithm~\ref{alg:cop}.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{dcdhscrs.png}
+    \caption{CRS Score Distribution}\label{fig:distcdcrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{dcdhsdrs.png}
+    \caption{DRS Score Distribution}\label{fig:distcddrs}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{icdhscrs.png}
+    \caption{CRS Iterations Distribution}\label{fig:itercdcrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{icdhsdrs.png}
+    \caption{DRS Iterations Distribution}\label{fig:itercddrs}
+  \end{subfigure}
+  \caption{Plots for the Cobb-Douglas Habitability Score.}\label{fig:cdhs}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{dceesacrs.png}
+    \caption{CRS Score Distribution}\label{fig:distcecrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{dceesadrs.png}
+    \caption{DRS Score Distribution}\label{fig:distcedrs}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{iceesacrs.png}
+    \caption{CRS Iterations Distribution}\label{fig:itercecrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{iceesadrs.png}
+    \caption{DRS Iterations Distribution}\label{fig:itercedrs}
+  \end{subfigure}
+  \caption{Plots for the Constant Elasticity Earth Similarity Approach.}\label{fig:ceesa}
+\end{figure}
 
 \begin{table}
+  \centering
+  \begin{subtable}{\textwidth}
     \centering
@@
     \end{tabular}
-  \caption{Estimated values of CDHS under CRS by PSO.\\
-    \footnotesize $\alpha$, $\beta$, $\gamma$ and $\delta$ record the parameters of Equation~\ref{eq:cdhscrs}. $Y_i$
-    and $Y_s$ record the maxima for the objective functions. $i_i$ and $i_s$ specify the number of iterations taken to
-    converge to a stable $\gbest$ value. Under the Class column there are four categories for the planets ---
-    Psychroplanets (psy), Mesoplanets (mes), Non-Habitable planets (non) and Hypopsychroplanets (hyp).  $\mathit{CDHS}$
-    is the final scores with $w_i=0.99$ and $w_s=0.01$.
-  }\label{tab:cdhscrs}
-\end{table}
+    \caption{Estimated habitability scores under CRS.}\label{tab:cdhscrs}
+  \end{subtable}
 
-\begin{table}
+  \vspace{4em}
+
+  \begin{subtable}{\textwidth}
     \centering
@@
       \midrule
-    GJ 176 b & non & 0.395 & 0.604 & 1.90 & 059 & 0.372 & 0.627 & 2.11 & 056 & 1.90\\
-    GJ 667 C b & non & 0.781 & 0.218 & 1.71 & 058 & 0.902 & 0.097 & 1.81 & 057 & 1.71\\
-    GJ 667 C e & psy & 0.179 & 0.820 & 1.40 & 049 & 0.234 & 0.765 & 1.39 & 060 & 1.40\\
-    GJ 667 C f & psy & 0.704 & 0.295 & 1.40 & 064 & 0.398 & 0.601 & 1.39 & 061 & 1.40\\
-    GJ 3634 b & non & 0.602 & 0.397 & 1.89 & 059 & 0.429 & 0.570 & 2.09 & 077 & 1.89\\
-    HD 20794 c & non & 0.014 & 0.985 & 1.35 & 050 & 0.116 & 0.883 & 1.34 & 045 & 1.35\\
-    HD 40307 e & non & 0.752 & 0.247 & 1.50 & 060 & 0.677 & 0.322 & 1.53 & 050 & 1.50\\
-    HD 40307 f & non & 0.887 & 0.112 & 1.52 & 051 & 0.261 & 0.738 & 1.55 & 060 & 1.52\\
-    HD 40307 g & psy & 0.300 & 0.699 & 1.82 & 062 & 0.785 & 0.214 & 1.98 & 056 & 1.82\\
-    Kepler-186 f & hyp & 0.073 & 0.926 & 1.17 & 046 & 0.740 & 0.259 & 1.12 & 051 & 1.17\\
-    Proxima Cen b & psy & 0.045 & 0.954 & 1.12 & 057 & 0.216 & 0.783 & 1.07 & 053 & 1.12\\
-    TRAPPIST-1 b & non & 0.102 & 0.897 & 1.09 & 041 & 0.000 & 0.000 & 1.00 & 065 & 1.09\\
-    TRAPPIST-1 c & non & 0.471 & 0.528 & 1.06 & 044 & 0.227 & 0.772 & 1.14 & 057 & 1.06\\
-    TRAPPIST-1 d & mes & 0.000 & 0.000 & 1.00 & 067 & 0.000 & 0.000 & 1.00 & 059 & 1.00\\
-    TRAPPIST-1 e & psy & 0.000 & 0.000 & 1.00 & 055 & 0.000 & 0.000 & 1.00 & 057 & 1.00\\
-    TRAPPIST-1 g & hyp & 0.888 & 0.111 & 1.13 & 047 & 0.949 & 0.050 & 1.09 & 046 & 1.13\\
+      GJ 176 b & non & 0.395 & 0.604 & 1.90 & 59 & 0.372 & 0.627 & 2.11 & 56 & 1.90\\
+      GJ 667 C b & non & 0.781 & 0.218 & 1.71 & 58 & 0.902 & 0.097 & 1.81 & 57 & 1.71\\
+      GJ 667 C e & psy & 0.179 & 0.820 & 1.40 & 49 & 0.234 & 0.765 & 1.39 & 60 & 1.40\\
+      GJ 667 C f & psy & 0.704 & 0.295 & 1.40 & 64 & 0.398 & 0.601 & 1.39 & 61 & 1.40\\
+      GJ 3634 b & non & 0.602 & 0.397 & 1.89 & 59 & 0.429 & 0.570 & 2.09 & 77 & 1.89\\
+      HD 20794 c & non & 0.014 & 0.985 & 1.35 & 50 & 0.116 & 0.883 & 1.34 & 45 & 1.35\\
+      HD 40307 e & non & 0.752 & 0.247 & 1.50 & 60 & 0.677 & 0.322 & 1.53 & 50 & 1.50\\
+      HD 40307 f & non & 0.887 & 0.112 & 1.52 & 51 & 0.261 & 0.738 & 1.55 & 60 & 1.52\\
+      HD 40307 g & psy & 0.300 & 0.699 & 1.82 & 62 & 0.785 & 0.214 & 1.98 & 56 & 1.82\\
+      Kepler-186 f & hyp & 0.073 & 0.926 & 1.17 & 46 & 0.740 & 0.259 & 1.12 & 51 & 1.17\\
+      Proxima Cen b & psy & 0.045 & 0.954 & 1.12 & 57 & 0.216 & 0.783 & 1.07 & 53 & 1.12\\
+      TRAPPIST-1 b & non & 0.102 & 0.897 & 1.09 & 41 & 0.000 & 0.000 & 1.00 & 65 & 1.09\\
+      TRAPPIST-1 c & non & 0.471 & 0.528 & 1.06 & 44 & 0.227 & 0.772 & 1.14 & 57 & 1.06\\
+      TRAPPIST-1 d & mes & 0.000 & 0.000 & 1.00 & 67 & 0.000 & 0.000 & 1.00 & 59 & 1.00\\
+      TRAPPIST-1 e & psy & 0.000 & 0.000 & 1.00 & 55 & 0.000 & 0.000 & 1.00 & 57 & 1.00\\
+      TRAPPIST-1 g & hyp & 0.888 & 0.111 & 1.13 & 47 & 0.949 & 0.050 & 1.09 & 46 & 1.13\\
       \bottomrule
     \end{tabular}
-  \caption{Estimated values of CDHS under DRS by PSO.\\
-    \footnotesize Here, $\alpha$, $\beta$, $\gamma$ and $\delta$ record the parameters of Equation~\ref{eq:cdhsdrs}. The
-    other columns are the same as those in Table~\ref{tab:cdhscrs}.
-  }\label{tab:cdhsdrs}
+    \caption{Estimated habitability scores under DRS.}\label{tab:cdhsdrs}
+  \end{subtable}
+  \caption{Cobb-Douglas Habitability Scores as estimated by \pso.\\
+    \footnotesize $\alpha$, $\beta$, $\gamma$ and $\delta$ record the parameters of Equation~\ref{eq:cdhscrs} in
+    Table~\ref{tab:cdhscrs} and the parameters of Equation~\ref{eq:cdhsdrs} in Table~\ref{tab:cdhsdrs}. $Y_i$ and $Y_s$
+    record the maxima for the objective functions. $i_i$ and $i_s$ specify the number of iterations taken to converge to
+    a stable $\gbest$ value. Under the Class column there are four categories for the planets --- Psychroplanets (psy),
+    Mesoplanets (mes), Non-Habitable planets (non) and Hypopsychroplanets (hyp).  $\mathit{CDHS}$ is the final score
+    with $w_i=0.99$ and $w_s=0.01$.
+  }
 \end{table}
@@
 \begin{table}
+  \centering
+  \begin{subtable}{\textwidth}
     \centering
@@
     \end{tabular}
-  \caption{Estimated values of CEESA under DRS by PSO.\\
-    \footnotesize $r$, $d$, $t$, $v$, $e$, $\rho$ and $eta$ record the parameters of Equation~\ref{eq:ceesadrs}.
-    $\mathit{CEESA}$ records the maxima of the objective function. $i$ specifies the number of iterations taken to
-    converge to the maxima.
-  }\label{tab:ceesadrs}
-\end{table}
+    \caption{Estimated habitability scores under DRS.}\label{tab:ceesadrs}
+  \end{subtable}
 
-\begin{table}
+  \vspace{4em}
+
+  \begin{subtable}{\textwidth}
     \centering
@@
     \end{tabular}
-  \caption{Estimated values of CEESA under CRS by PSO.\\
-    \footnotesize Here, $r$, $d$, $t$, $v$, $e$ and $\rho$ record the parameters of Equation~\ref{eq:ceesacrs}. Since
-    under the CRS constraint $\eta=1$, there is no need for the parameter in the problem.
-  }\label{tab:ceesacrs}
+    \caption{Estimated habitability scores under CRS.}\label{tab:ceesacrs}
+  \end{subtable}
+  \caption{Constant Elasticity Earth Similarity Approach scores as estimated by \pso.\\
+      \footnotesize $r$, $d$, $t$, $v$, $e$, $\rho$ and $eta$ record the parameters of Equation~\ref{eq:ceesadrs} in
+      Table~\ref{tab:ceesadrs} and the parameters of Equation~\ref{eq:ceesacrs} in Table~\ref{tab:ceesacrs}. However,
+      since under the CRS constraint, $\eta=1$, there is no need for the parameter $\eta$ in the problem. The column
+      $\mathit{CEESA}$ records the maxima of the objective function. $i$ specifies the number of iterations taken to
+      converge to the maxima.
+    }
 \end{table}
-
-% \begin{pointers}
-% \item The values and proximity to earth's habitability score.
-% \item Values that do not converge. Or were hard to converge.
-% \item Speed of convergence graphs.
-% \item Graph1: Iterations to convergence vs. Number of particles.
-% \item Graph2: Distribution of number of iterations to convergence.
-% \item Graph3: Iterations to convergence vs. Constraint parameters.
-% \end{pointers}
 
@@
 
+% Include a description of max_iter and stable_iter.
 Matrices $P$ and $V$ represent current position and velocity, where the $i$\textsuperscript{th} row of each correspond

### 2018-01-17 docs/report/draft.tex
Finished running time tests.

@@
 
-The plots in Figures~\ref{fig:distcdcrs} and~\ref{fig:distcddrs} indicate the distribution of the CDHS across the
-exoplanets tested from the PHL-EC.\@ Figures~\ref{fig:itercdcrs} and~\ref{fig:itercddrs} show the distribution of
-iterations required to converge to a global maxima. The spike at 0 is caused by particles converging to a $\gbest$ that
-does not shift from the original position (for a more detailed explanation see Appendix \ref{app:con}). The plots in
-Figures~\ref{fig:distcedrs} and~\ref{fig:distcedrs} indicate the distribution of the CEESA score across the exoplanets,
-while Figures~\ref{fig:itercedrs} and~\ref{fig:itercecrs} indicate the distribution of iterations to convergence. These
-graphs aggregate the results of optimizing the Habitability Production Functions (Equations~\ref{eq:cdhscrs},
-~\ref{eq:cdhsdrs},~\ref{eq:ceesadrs} and~\ref{eq:ceesacrs}) for each exoplanet in the PHL-EC by the method described in
-Algorithm~\ref{alg:cop}.
+The plots in Figures~\ref{fig:distcdcrs} and~\ref{fig:distcddrs} describe the distribution of the CDHS across exoplanets
+tested from the PHL-EC.\@ Figures~\ref{fig:itercdcrs} and~\ref{fig:itercddrs} show the distribution of iterations
+required to converge to a global maxima. The spike at 0 is caused by particles converging to a $\gbest$ that does not
+shift from the original position (for a more detailed explanation see Appendix~\ref{app:con}). The plots in
+Figures~\ref{fig:distcedrs} and~\ref{fig:distcedrs} describe the distribution of the CEESA score across the exoplanets,
+while Figures~\ref{fig:itercedrs} and~\ref{fig:itercecrs} show the distribution of iterations to convergence. These
+graphs aggregate the results of optimizing the Habitability Production Functions
+(Equations~\ref{eq:cdhscrs},~\ref{eq:cdhsdrs},~\ref{eq:ceesadrs} and~\ref{eq:ceesacrs}) for each exoplanet in the PHL-EC
+by the method described in Algorithm~\ref{alg:cop}.
 

### 2018-01-17 source/ceesa/ceesa.py

@@ -13,7 +13,8 @@ MESSAGE = '{:25}{:>7.5}{:8.4f}{:8.4f}{:8.4f}{:8.4f}{:8.4f}'\
         '{:8.4f}{:8.4f}{:10.4f}{:6}'
 TITLE = '{:25}{:>7}{:>8}{:>8}{:>8}{:>8}{:>8}{:>8}{:>8}{:>10}{:>6}'
 ERROR = '{:25}{:^79}'
-HEADERS = ('Name', 'Cls', 'r', 'd', 't', 'v', 'e', 'Rho', 'Eta', 'CEESA', 'Iter')
+HEADERS = ('Name', 'Cls', 'r', 'd', 't', 'v', 'e',
+           'Rho', 'Eta', 'CEESA', 'Iter')
 ERR_TEXT = '** Convergence failed. **'
 TOTAL_CHAR = 104
 PROGRESS_BAR = '[{:' + str(TOTAL_CHAR - 10) + '}]  ({:>3}%)'

### 2018-01-17 docs/report/draft.tex
Tables are now autogenerated and loaded.

@@
-\documentclass[9pt]{article}
+\documentclass[10pt]{article}
 
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=1.2in]{geometry}
 \usepackage[utf8]{inputenc}
@@
 \graphicspath{{figs/}}
-\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
+% \sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
 
@@
 \date{\today}
-
-% Useful for initial drafts.
-\newenvironment{pointers}{%
-  \noindent Should include,
-  \begin{itemize}
-    \setlength{\itemsep}{-1pt}}{%
-\end{itemize}}
 
@@
 
-The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
-Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL). The catalog records observed and
-modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param} describes the
-parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and eccentricity
-are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface temperature (288K) and
-eccentricity (0.017). PHL-EC assumes an Eccentricity of 0 when unavailable. The PHL-EC records empty values for planets
-whose surface temperature is not known. We chose to drop these records from the experiment.
-
 \begin{table}
@@
 
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{dcdhscrs.png}
+    \caption{CRS Score Distribution}\label{fig:distcdcrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{dcdhsdrs.png}
+    \caption{DRS Score Distribution}\label{fig:distcddrs}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{icdhscrs.png}
+    \caption{CRS Iterations Distribution}\label{fig:itercdcrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{icdhsdrs.png}
+    \caption{DRS Iterations Distribution}\label{fig:itercddrs}
+  \end{subfigure}
+  \caption{Plots for the Cobb-Douglas Habitability Score.}\label{fig:cdhs}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{dceesacrs.png}
+    \caption{CRS Score Distribution}\label{fig:distcecrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{dceesadrs.png}
+    \caption{DRS Score Distribution}\label{fig:distcedrs}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{iceesacrs.png}
+    \caption{CRS Iterations Distribution}\label{fig:itercecrs}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[b]{0.38\textwidth}
+    \includegraphics[width=\textwidth]{iceesadrs.png}
+    \caption{DRS Iterations Distribution}\label{fig:itercedrs}
+  \end{subfigure}
+  \caption{Plots for the Constant Elasticity Earth Similarity Approach.}\label{fig:ceesa}
+\end{figure}
+
+The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
+Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL). The catalog records observed and
+modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param} describes the
+parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and eccentricity
+are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface temperature
+(\SI{288}{\kelvin}) and eccentricity (\num{0.017}). PHL-EC assumes an Eccentricity of 0 when unavailable. The PHL-EC
+records empty values for planets whose surface temperature is not known. We chose to drop these records from the
+experiment.
+
 Our implemenation used $n=25$ particles to traverse the search space, with learning rates $\lambda_g=0.8$ and
@@
 details are discussed in Appendix~\ref{app:imp}.
-
-Tables~\ref{tab:cdhscrs} and~\ref{tab:cdhsdrs} record the CDHS values for a sample of exoplanets under CRS and DRS
-respectively at $w_i=0.99$ and $w_s=0.01$. Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the
-CEESA scores. All tables also record the number of iterations taken to converge to a stable $\gbest$ value. As seen in
-the tables, although CEESA has 7 parameters and 16 constraints under DRS, PSO takes just over twice the number of
-iterations to converge as in each step of CDHS, which has 2 parameters and 5 constraints. This is a promising result as
-it indicates that the iterations required for converging increases sub-linearly with the number of parameters in the
-model.
 
@@
 
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{dcdhscrs.png}
-    \caption{CRS Score Distribution}\label{fig:distcdcrs}
-  \end{subfigure}
-  \quad
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{dcdhsdrs.png}
-    \caption{DRS Score Distribution}\label{fig:distcddrs}
-  \end{subfigure}
+Tables~\ref{tab:cdhscrs} and~\ref{tab:cdhsdrs} record the CDHS values for a sample of exoplanets under CRS and DRS
+respectively at $w_i=0.99$ and $w_s=0.01$. Tables~\ref{tab:ceesacrs} and~\ref{tab:ceesadrs} record the same for the
+CEESA scores. All tables also record the number of iterations taken to converge to a stable $\gbest$ value. As seen in
+the tables, although CEESA has 7 parameters and 16 constraints under DRS, PSO takes just over twice the number of
+iterations to converge as in each step of CDHS, which has 2 parameters and 5 constraints. This is a promising result as
+it indicates that the iterations required for converging increases sub-linearly with the number of parameters in the
+model.
 
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{icdhscrs.png}
-    \caption{CRS Iterations Distribution}\label{fig:itercdcrs}
-  \end{subfigure}
-  \quad
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{icdhsdrs.png}
-    \caption{DRS Iterations Distribution}\label{fig:itercddrs}
-  \end{subfigure}
-  \caption{Plots for the Cobb-Douglas Habitability Score.}\label{fig:cdhs}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{dceesacrs.png}
-    \caption{CRS Score Distribution}\label{fig:distcecrs}
-  \end{subfigure}
-  \quad
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{dceesadrs.png}
-    \caption{DRS Score Distribution}\label{fig:distcedrs}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{iceesacrs.png}
-    \caption{CRS Iterations Distribution}\label{fig:itercecrs}
-  \end{subfigure}
-  \quad
-  \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{iceesadrs.png}
-    \caption{DRS Iterations Distribution}\label{fig:itercedrs}
-  \end{subfigure}
-  \caption{Plots for the Constant Elasticity Earth Similarity Approach.}\label{fig:ceesa}
-\end{figure}
+As for real time taken to converge, PSO took \SI{666.85}{\second} ($\approx\SI{11}{\minute}\ \SI{7}{\second}$) to
+estimate the CDHS under CRS for \num{1683} exoplanets, at an average of \SI{198.11}{\milli\second} for each planet for
+each part of the CDHS.\@ For CDHS under DRS, it took \SI{638.69}{\second} ($\approx\SI{10}{\minute}\ \SI{39}{\second}$)
+at an average of \SI{189.75}{\milli\second} for each part of the CDHS.\@ The CEESA estimates, requiring one calculation,
+took a little over half the total time to execute. Under DRS it took a total of \SI{370.86}{\second}
+($\approx\SI{6}{\minute}\ \SI{11}{\second}$) at \SI{220.36}{\milli\second} per planet, while under CRS it took
+\SI{356.92}{\second} ($\approx\SI{5}{\minute}\ \SI{57}{\second}$) at \SI{212.07}{\milli\second} per planet. This is
+considerably fast considering the number of parameters and constraints for the problem.
 
@@
     \centering
-    \begin{tabular}{l r r r r r r r r r r}
-      \toprule
-      Name & Class & $\alpha$ & $\beta$ & $Y_i$ & $i_i$ & $\gamma$ & $\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\
-      \midrule
-      GJ 176 b & non & 0.460 & 0.540 & 1.90 & 50 & 0.107 & 0.893 & 2.11 & 61 & 1.90\\
-      GJ 667 C b & non & 0.423 & 0.577 & 1.71 & 58 & 0.692 & 0.308 & 1.81 & 54 & 1.71\\
-      GJ 667 C e & psy & 0.129 & 0.871 & 1.40 & 50 & 0.258 & 0.742 & 1.39 & 55 & 1.40\\
-      GJ 667 C f & psy & 0.534 & 0.466 & 1.40 & 48 & 0.865 & 0.135 & 1.39 & 47 & 1.40\\
-      GJ 3634 b & non & 0.409 & 0.591 & 1.89 & 58 & 0.724 & 0.276 & 2.09 & 48 & 1.89\\
-      HD 20794 c & non & 0.260 & 0.740 & 1.35 & 50 & 0.096 & 0.904 & 1.34 & 58 & 1.35\\
-      HD 40307 e & non & 0.168 & 0.832 & 1.50 & 49 & 0.636 & 0.364 & 1.53 & 63 & 1.50\\
-      HD 40307 f & non & 0.702 & 0.298 & 1.52 & 68 & 0.303 & 0.697 & 1.55 & 45 & 1.52\\
-      HD 40307 g & psy & 0.964 & 0.036 & 1.82 & 51 & 0.083 & 0.917 & 1.98 & 55 & 1.82\\
-      Kepler-186 f & hyp & 0.338 & 0.662 & 1.17 & 50 & 0.979 & 0.021 & 1.12 & 40 & 1.17\\
-      Proxima Cen b & psy & 0.515 & 0.484 & 1.12 & 37 & 0.755 & 0.245 & 1.07 & 00 & 1.12\\
-      TRAPPIST-1 b & non & 0.319 & 0.681 & 1.09 & 00 & 0.801 & 0.199 & 0.89 & 00 & 1.09\\
-      TRAPPIST-1 c & non & 0.465 & 0.535 & 1.06 & 00 & 0.935 & 0.065 & 1.14 & 26 & 1.06\\
-      TRAPPIST-1 d & mes & 0.635 & 0.365 & 0.77 & 34 & 0.475 & 0.525 & 0.73 & 47 & 0.77\\
-      TRAPPIST-1 e & psy & 0.145 & 0.855 & 0.92 & 00 & 0.897 & 0.103 & 0.83 & 55 & 0.92\\
-      TRAPPIST-1 g & hyp & 0.226 & 0.774 & 1.13 & 43 & 0.876 & 0.124 & 1.09 & 00 & 1.13\\
-      \bottomrule
-    \end{tabular}
+    \include{./tabs/cdhscrs}
     \caption{Estimated habitability scores under CRS.}\label{tab:cdhscrs}
@@
     \centering
-    \begin{tabular}{l r r r r r r r r r r}
-      \toprule
-      Name & Class & $\alpha$ & $\beta$ & $Y_i$ & $i_i$ & $\gamma$ & $\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\
-      \midrule
-      GJ 176 b & non & 0.395 & 0.604 & 1.90 & 59 & 0.372 & 0.627 & 2.11 & 56 & 1.90\\
-      GJ 667 C b & non & 0.781 & 0.218 & 1.71 & 58 & 0.902 & 0.097 & 1.81 & 57 & 1.71\\
-      GJ 667 C e & psy & 0.179 & 0.820 & 1.40 & 49 & 0.234 & 0.765 & 1.39 & 60 & 1.40\\
-      GJ 667 C f & psy & 0.704 & 0.295 & 1.40 & 64 & 0.398 & 0.601 & 1.39 & 61 & 1.40\\
-      GJ 3634 b & non & 0.602 & 0.397 & 1.89 & 59 & 0.429 & 0.570 & 2.09 & 77 & 1.89\\
-      HD 20794 c & non & 0.014 & 0.985 & 1.35 & 50 & 0.116 & 0.883 & 1.34 & 45 & 1.35\\
-      HD 40307 e & non & 0.752 & 0.247 & 1.50 & 60 & 0.677 & 0.322 & 1.53 & 50 & 1.50\\
-      HD 40307 f & non & 0.887 & 0.112 & 1.52 & 51 & 0.261 & 0.738 & 1.55 & 60 & 1.52\\
-      HD 40307 g & psy & 0.300 & 0.699 & 1.82 & 62 & 0.785 & 0.214 & 1.98 & 56 & 1.82\\
-      Kepler-186 f & hyp & 0.073 & 0.926 & 1.17 & 46 & 0.740 & 0.259 & 1.12 & 51 & 1.17\\
-      Proxima Cen b & psy & 0.045 & 0.954 & 1.12 & 57 & 0.216 & 0.783 & 1.07 & 53 & 1.12\\
-      TRAPPIST-1 b & non & 0.102 & 0.897 & 1.09 & 41 & 0.000 & 0.000 & 1.00 & 65 & 1.09\\
-      TRAPPIST-1 c & non & 0.471 & 0.528 & 1.06 & 44 & 0.227 & 0.772 & 1.14 & 57 & 1.06\\
-      TRAPPIST-1 d & mes & 0.000 & 0.000 & 1.00 & 67 & 0.000 & 0.000 & 1.00 & 59 & 1.00\\
-      TRAPPIST-1 e & psy & 0.000 & 0.000 & 1.00 & 55 & 0.000 & 0.000 & 1.00 & 57 & 1.00\\
-      TRAPPIST-1 g & hyp & 0.888 & 0.111 & 1.13 & 47 & 0.949 & 0.050 & 1.09 & 46 & 1.13\\
-      \bottomrule
-    \end{tabular}
+    \include{./tabs/cdhsdrs}
     \caption{Estimated habitability scores under DRS.}\label{tab:cdhsdrs}
@@
     \centering
-    \begin{tabular}{l r r r r r r r r r r}
-      \toprule
-      Name & Class & $r$ & $d$ & $t$ & $v$ & $e$ & $\rho$ & $\eta$ & $\mathit{CEESA}$ & $i$\\
-      \midrule
-      GJ 176 b & non & 0.304 & 0.001 & 0.375 & 0.271 & 0.050 & 0.467 & 0.808 & 1.52 &  85\\
-      GJ 667 C b & non & 0.297 & 0.010 & 0.318 & 0.052 & 0.322 & 0.682 & 0.730 & 2.36 &  90\\
-      GJ 667 C e & psy & 0.230 & 0.286 & 0.137 & 0.199 & 0.148 & 0.551 & 0.906 & 1.14 &  85\\
-      GJ 667 C f & psy & 0.397 & 0.035 & 0.152 & 0.402 & 0.014 & 0.793 & 0.999 & 1.31 & 100\\
-      GJ 3634 b & non & 0.178 & 0.175 & 0.005 & 0.194 & 0.447 & 0.894 & 0.657 & 2.07 &  94\\
-      HD 20794 c & non & 0.073 & 0.142 & 0.452 & 0.190 & 0.144 & 0.953 & 0.635 & 1.20 &  78\\
-      HD 40307 e & non & 0.156 & 0.307 & 0.185 & 0.033 & 0.319 & 0.428 & 0.939 & 2.69 &  88\\
-      HD 40307 f & non & 0.272 & 0.231 & 0.064 & 0.305 & 0.127 & 0.676 & 0.802 & 1.28 &  77\\
-      HD 40307 g & psy & 0.113 & 0.219 & 0.066 & 0.454 & 0.148 & 0.711 & 0.991 & 3.26 &  92\\
-      Kepler-186 f & hyp & 0.039 & 0.159 & 0.116 & 0.329 & 0.357 & 0.253 & 0.919 & 1.35 &  70\\
-      Proxima Cen b & psy & 0.272 & 0.173 & 0.284 & 0.193 & 0.079 & 0.615 & 0.114 & 0.99 &  75\\
-      TRAPPIST-1 b & non & 0.488 & 0.151 & 0.039 & 0.193 & 0.129 & 0.151 & 0.014 & 0.99 &  87\\
-      TRAPPIST-1 c & non & 0.172 & 0.236 & 0.275 & 0.242 & 0.075 & 0.969 & 0.962 & 1.06 &  80\\
-      TRAPPIST-1 d & mes & 0.106 & 0.308 & 0.075 & 0.218 & 0.293 & 0.844 & 0.017 & 0.99 &  93\\
-      TRAPPIST-1 e & psy & 0.189 & 0.266 & 0.192 & 0.094 & 0.260 & 0.371 & 0.006 & 0.99 &  84\\
-      TRAPPIST-1 g & hyp & 0.326 & 0.186 & 0.143 & 0.278 & 0.067 & 0.315 & 0.021 & 1.00 &  76\\
-      \bottomrule
-    \end{tabular}
+    \include{./tabs/ceesadrs}
     \caption{Estimated habitability scores under DRS.}\label{tab:ceesadrs}
@@
     \centering
-    \begin{tabular}{l r r r r r r r r r}
-      \toprule
-      Name & Class & $r$ & $d$ & $t$ & $v$ & $e$ & $\rho$ & $\mathit{CEESA}$ & $i$\\
-      \midrule
-      GJ 176 b & non & 0.194 & 0.020 & 0.315 & 0.465 & 0.006 & 0.398 & 1.88 &  86\\
-      GJ 667 C b & non & 0.162 & 0.289 & 0.090 & 0.087 & 0.372 & 0.836 & 3.54 & 107\\
-      GJ 667 C e & psy & 0.373 & 0.032 & 0.134 & 0.304 & 0.157 & 0.217 & 1.25 &  71\\
-      GJ 667 C f & psy & 0.394 & 0.006 & 0.043 & 0.360 & 0.196 & 0.490 & 1.44 &  81\\
-      GJ 3634 b & non & 0.351 & 0.122 & 0.006 & 0.069 & 0.453 & 0.439 & 2.89 &  96\\
-      HD 20794 c & non & 0.101 & 0.077 & 0.691 & 0.071 & 0.059 & 0.756 & 1.58 &  94\\
-      HD 40307 e & non & 0.069 & 0.091 & 0.097 & 0.173 & 0.569 & 0.768 & 5.29 &  94\\
-      HD 40307 f & non & 0.285 & 0.161 & 0.053 & 0.443 & 0.058 & 0.342 & 1.42 &  73\\
-      HD 40307 g & psy & 0.156 & 0.010 & 0.081 & 0.302 & 0.451 & 0.612 & 7.15 &  94\\
-      Kepler-186 f & hyp & 0.036 & 0.017 & 0.082 & 0.383 & 0.483 & 0.929 & 1.68 &  85\\
-      Proxima Cen b & psy & 0.352 & 0.383 & 0.103 & 0.059 & 0.103 & 0.936 & 0.89 &  83\\
-      TRAPPIST-1 b & non & 0.148 & 0.147 & 0.344 & 0.269 & 0.093 & 0.767 & 0.94 &  81\\
-      TRAPPIST-1 c & non & 0.038 & 0.060 & 0.575 & 0.321 & 0.005 & 0.602 & 1.17 &  86\\
-      TRAPPIST-1 d & mes & 0.023 & 0.065 & 0.475 & 0.391 & 0.045 & 0.830 & 0.84 &  79\\
-      TRAPPIST-1 e & psy & 0.176 & 0.464 & 0.253 & 0.103 & 0.004 & 0.920 & 0.86 &  81\\
-      TRAPPIST-1 g & hyp & 0.060 & 0.086 & 0.310 & 0.540 & 0.004 & 0.848 & 0.97 &  86\\
-      \bottomrule
-    \end{tabular}
+    \include{./tabs/ceesacrs}
     \caption{Estimated habitability scores under CRS.}\label{tab:ceesacrs}
@@
 \section{Conclusions}
-\begin{pointers}
-\item Why is the speed so important?
-\item Parallelizable.
-\end{pointers}
+% - Why is the speed so important?
+% - Parallelizable.
 

### 2018-01-17 extract_sample.py

@@ -32,15 +32,17 @@ planets = [
         'TRAPPIST-1 g',
         ]
 
-with open('cdhs_sample_table.txt', 'w') as f:
-    def fprint(*args, **kwargs): print(*args, **kwargs, file=f)
-    for c, d in cdhs.items():
-        fprint(c.upper() + ':')
-        fprint('\\begin{tabular}{l l r r r r r r r r r}')
-        fprint('\\toprule')
-        fprint('Name & Class & $\\alpha$ & $\\beta$ & $Y_i$ & $i_i$ & '
+
+for c, d in cdhs.items():
+    tabline = '  {} & {} & ${:.3f}$ & ${:.3f}$ & ${:.2f}$ & ${:2}$ & ${:.3f}$'\
+              ' & ${:.3f}$ & ${:.2f}$ & ${:2}$ & ${:.2f}$\\\\'
+    with open('docs/report/tabs/cdhs%s.tex' % (c,), 'w') as f:
+        def fprint(*args, **kwargs): print(*args, **kwargs, file=f)
+        fprint('\\begin{tabular}{l r r r r r r r r r r}')
+        fprint('  \\toprule')
+        fprint('  Name & Class & $\\alpha$ & $\\beta$ & $Y_i$ & $i_i$ & '
                '$\\gamma$ & $\\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\\\')
-        fprint('\\midrule')
+        fprint('  \\midrule')
         for _, row in d[d.Name.isin(planets)].iterrows():
             habc = row['Cls'][:3]
             vals = np.round(np.array([*(row[2:])]), 4)
@@ -49,28 +51,27 @@ with open('cdhs_sample_table.txt', 'w') as f:
                 vals[1] /= 1.001
                 vals[3] /= 1.001
                 vals[4] /= 1.001
-            fprint('{} & {} & {:.3f} & {:.3f} & {:.2f} & {:2} & {:.3f} & '
-                   '{:.3f} & {:.2f} & {:2} & {:.2f}\\\\'.format(
-                       row['Name'], habc,
-                       vals[0], vals[1], vals[2], int(vals[7]),
-                       vals[3], vals[4], vals[5], int(vals[8]), vals[6]))
-        fprint('\\bottomrule')
+            fprint(tabline.format(row['Name'], habc, *vals[:3], int(vals[7]),
[…]

### 2018-01-18 docs/report/draft.tex
Finished report.

@@
-\documentclass[10pt]{article}
+\documentclass[10pt,draft]{article}
 
+\usepackage[
+  citestyle=numeric-comp,
+  sorting=none]{biblatex}
 \usepackage[margin=1.2in]{geometry}
@@
 
+\usepackage[title]{appendix}
 \usepackage{booktabs}
@@
 \graphicspath{{figs/}}
-% \sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
+\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
 
@@
 \newcommand{\pso}{Particle Swarm Optimization}
+\newcommand{\tsup}{\textsuperscript}
 
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \DeclareMathOperator*{\sgn}{sgn}
@@
 
+% Bibliography
+\addbibresource{bibliography.bib}
+
 \begin{document}
@@
 
-The search for extra-terrestrial life and potentially habitable extrasolar planets has been an international venture
-demanding large investments in cost and effort, since Frank Drake's attempt with Project Ozma in the mid-20th century.
-The first exoplanet was officially confirmed in 1992 which marked the start of a trend that has lasted 25 years and
-yielded over 3,700 confirmed exoplanets. There have been attempts to assess the habitability of these planets and to
-assign a score based on their similarity to Earth. Two such habitability scores are the Cobb-Douglas Habitability Score
-(CDHS) and the Constant Elasticity Earth Similarity Approach (CEESA) score. Estimating these scores involves maximizing
-a production function while observing a set of constraints on the input variables.
+The search for extra-terrestrial life~\cite{Seth1,Biosig} and potentially habitable extrasolar
+planets~\cite{Kepler,Presto} has been an international venture since Frank Drake's attempt with Project Ozma in the
+mid-20th century~\cite{Ozma}, demanding large investments in cost and effort. \textcite{Exo1} confirmed the first
+exoplanet in 1991. This marked the start of a trend that has lasted 25 years and yielded over 3,700 confirmed
+exoplanets. There have been attempts to assess the habitability of these planets and to assign a score based on their
+similarity to Earth. Two such habitability scores are the Cobb-Douglas Habitability Score (CDHS)~\cite{Bora,Saha} and
+the Constant Elasticity Earth Similarity Approach (CEESA) score. Estimating these scores involves maximizing a
+production function while observing a set of constraints on the input variables.
 
@@
 optimizing a continuous function that does away with the need for a gradient. It employs a large number of particles
-that traverse the search space converging toward a global best solution encountered by at least one of the particles.
+that traverse the search space converging toward a global best solution encountered by at least one of the
+particles~\cite{PSO1,PSO2}.
 
@@
 through variations in initializing and updating particles constraints are straightforward to represent and adhere to, as
-seen in Section~\ref{subsec:copso}.
-
-% Pull out citations from the Applications paper.
-PSO has been adapted to a wide range of design optimization problems including network and VLSI design. It has found
-applications in machine learning under clustering, feature detection and classification. As a modeling paradigm, it has
-been used for constructing customer satisfaction models, modeling MIDI music and chaotic time series modeling.
+seen in Section~\ref{subsec:copso}. \textcite{Poli1,Poli2} carried out extensive surveys on the applications of PSO
+reporting uses in Communication Networks, Machine Learning, Design, Combinatorial Optimization and Modeling, among
+others.
 
 This paper demonstrates the applicability of \pso\ in estimating CDHS and CEESA scores of an exoplanet by maximizing
-their respective production functions, discussed in Sections~\ref{subsec:cdhs} and~\ref{subsec:ceesa}. CDHS considers
+their respective production functions (discussed in Sections~\ref{subsec:cdhs} and~\ref{subsec:ceesa}). CDHS considers
 the planet's Radius, Mass, Escape Velocity and Surface Temperature, while CEESA includes a fifth parameter, the Orbital
 Eccentricity of the planet. The Exoplanets Catalog hosted by the Planetary Habitability Laboratory, UPR Arecibo records
-these parameters for each exoplanet in Earth Units. Section~\ref{sec:results} reports the performance of PSO and
-discusses the distribution of the habitability scores of the exoplanets.
+these parameters for each exoplanet in Earth Units~\cite{PHL}. Section~\ref{sec:results} reports the performance of PSO
+and discusses the distribution of the habitability scores of the exoplanets.
 
 \section{Habitability Scores}\label{sec:habscore}
-% Add the citations to Saha's papers.
 
 \subsection{Cobb-Douglas Habitability Score}\label{subsec:cdhs}
-Estimating the Cobb-Douglas Habitability Score (CDHS) requires estimating an interior CDHS (CDHS\textsubscript{i}) and a
-surface CDHS (CDHS\textsubscript{s}) by maximizing the following production functions,
+Estimating the Cobb-Douglas Habitability Score (CDHS)~\cite{Bora} requires estimating an interior score \linebreak
+($\mathit{CDHS}_i$) and a surface score ($\mathit{CDHS}_s$) by maximizing the following production functions,
 \begin{subequations}
   \begin{alignat}{4}
-    Y_i\ &=\ {CDHS}_i\ &=&\ R^\alpha.&D^\beta\label{eq:cdhsi}\\
-    Y_s\ &=\ {CDHS}_s\ &=&\ {V_e}^\gamma.&{T_s}^\delta\label{eq:cdhss}
+    Y_i\ &=\ {CDHS}_i\ &=&\ R^\alpha.&D^\beta\,,\label{eq:cdhsi}\\
+    Y_s\ &=\ {CDHS}_s\ &=&\ {V_e}^\gamma.&{T_s}^\delta\,,\label{eq:cdhss}
   \end{alignat}
@@
 \begin{equation}
-  0 < \alpha,\beta,\gamma,\delta < 1
+  0 < \alpha,\beta,\gamma,\delta < 1\,.
 \end{equation}
@@
   \begin{alignat}{3}
-    \text{CRS:} & \quad\alpha+\beta = 1,&\quad\gamma+\delta = 1,\\
-    \text{DRS:} & \quad\alpha+\beta < 1,&\quad\gamma+\delta < 1.
+    \text{CRS:} & \quad\alpha+\beta = 1,&\quad\gamma+\delta = 1\,,\\
+    \text{DRS:} & \quad\alpha+\beta < 1,&\quad\gamma+\delta < 1\,.
   \end{alignat}
@@
 \begin{equation}
-  Y\ =\ w_i.Y_i + w_s.Y_s
+  Y\ =\ w_i.Y_i + w_s.Y_s\,.
 \end{equation}
@@
 \subsection{Constant Elasticity Earth Similarity Approach}\label{subsec:ceesa}
+% Cite what?
 The Constant Elasticity Earth Similarity Approach (CEESA) uses the following production function to estimate the
@@
 \begin{equation}\label{eq:ceesa}
-  Y = {(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\frac{\eta}{\rho}}
+  Y = {(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\frac{\eta}{\rho}}\,,
 \end{equation}
@@
   \begin{align}
-      0 < r,d,t,v,e < 1\\
-      r+d+t+v+e = 1
+      0 < r,d,t,v,e < 1\,,\\
+      r+d+t+v+e = 1\,.
   \end{align}
@@
   \begin{alignat}{3}
-    \text{CRS:} & \quad 0 < \eta \leq 1,\\
-    \text{DRS:} & \quad \eta = 1.
+    \text{CRS:} & \quad 0 < \eta \leq 1\,,\\
+    \text{DRS:} & \quad \eta = 1\,.
   \end{alignat}
@@
 
-Particle Swarm Optimization (PSO) is a biologically inspired metaheuristic for finding the global minima of a function.
-Traditionally designed for unconstrained inputs, it works by iteratively converging a population of randomly initialized
-solutions, called particles, toward a globally optimal solution. Each particle in the population keeps track of its
-current position and the best solution it has encountered so far, called $\pbest$. Each particle also has an associated
-randomized velocity used to traverse the search space. The swarm keeps track of the overall best solution, called
-$\gbest$. Each iteration of the swarm updates the velocity of the particle towards its $\pbest$ and the $\gbest$ values.
+% Cite: PSO paper.
+Particle Swarm Optimization (PSO)~\cite{PSO1} is a biologically inspired metaheuristic for finding the global minima of
+a function.  Traditionally designed for unconstrained inputs, it works by iteratively converging a population of
+randomly initialized solutions, called particles, toward a globally optimal solution. Each particle in the population
+keeps track of its current position and the best solution it has encountered so far, called $\pbest$. Each particle also
+has an associated randomized velocity used to traverse the search space. The swarm keeps track of the overall best
+solution, called $\gbest$. Each iteration of the swarm updates the velocity of the particle towards its $\pbest$ and the
+$\gbest$ values.
 
@@
 to its initial position. The $\pbest$ position that corresponds to the minimum fitness is selected to be the $\gbest$
-position of the swarm.
+position of the swarm. \textcite{PSO2} discussed the use of inertial weights to regulate velocity to balance the global
+and local search. Upper and lower bounds limit velocity within $\pm v_\mathit{max}$.
 
@@
 random numbers $u_g, u_p$ from a uniform distribution, $U(0,1)$ and updates the particle velocity as indicated in
-line~\ref{algline:vup}. Here, $\mu$ is the friction coefficient and $\lambda_g,\lambda_p$ are the global and particle
+line~\ref{algline:vup}. Here, $\omega$ is the inertial weight and $\lambda_g,\lambda_p$ are the global and particle
 learning rates. If the new position of the particle corresponds to a better fit than its $\pbest$, the algorithm updates
@@
         \State{$u_p, u_g \sim U(0,1)$}
-        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({\gbest-p_i}) + \lambda_p.u_p.({{\pbest}_i-p_i})$}\label{algline:vup}
+        \State{$v_i \gets \omega.v_i + \lambda_g.u_g.({\gbest-p_i}) + \lambda_p.u_p.({{\pbest}_i-p_i})$}\label{algline:vup}
+        \State{$v_i \gets \sgn(v_i).\max\{|v_\mathit{max}|, |v_i|\}$}
         \State{$p_i \gets p_i + v_i$}
@@
       \State{$\gbest \gets \smashoperator{\argmin\limits_{{\pbest}_i,\,i=1\dots n}} f({\pbest}_i)$}
-    \Until{$|\oldbest - \gbest| < threshold$}
+    \Until{$|\oldbest - \gbest| < \mathit{threshold}$}
     \State{\textbf{return} {$f(\gbest)$}}
@@
 solution. Solving the initialization problem is straightforward, resample each random solution from the uniform
-distribution until every initial solution is feasible. To solve the convergence problem, each particle uses another
+distribution until every initial solution is feasible. To solve the convergence problem each particle uses another
 particle's $\pbest$ value, called $\lbest$, instead of its own to update its velocity. Algorithm~\ref{alg:cop} describes
@@
 
-On each iteration, for each particle, the algorithm first picks two random numbers $u_g,u_p$ as before. It then selects
-a $\pbest$ value from all particles in the swarm that is closest to the position of the particle being updated as its
-$\lbest$. The $\lbest$ value substitutes ${\pbest}_i$ in the velocity update equation. While updating $\pbest$ for the
-particle, the algorithm checks if the current fit is better than $\pbest$, and performs the update if the current
-position satisfies all constraints. The algorithm updates $\gbest$ as before.
+On each iteration, for each particle, the algorithm first picks two random numbers $u_g,u_p$. It then selects a $\pbest$
+value from all particles in the swarm that is closest to the position of the particle being updated as its $\lbest$. The
+$\lbest$ value substitutes ${\pbest}_i$ in the velocity update equation. While updating $\pbest$ for the particle, the
+algorithm checks if the current fit is better than $\pbest$, and performs the update if the current position satisfies
+all constraints. The algorithm updates $\gbest$ as before.
 
@@
             p_i\|}^2$}\label{algline:lbest}
-        \State{$v_i \gets \mu.v_i + \lambda_g.u_g.({\gbest-p_i}) + \lambda_p.u_p.({\lbest-p_i})$}\label{algline:cvup}
+        \State{$v_i \gets \omega.v_i + \lambda_g.u_g.({\gbest-p_i}) + \lambda_p.u_p.({\lbest-p_i})$}\label{algline:cvup}
         \State{$p_i \gets p_i + v_i$}
@@
     & \text{subject to}
-    & & g_k(x) \leq 0,&\; k = 1\dots q,\\
-    &&& h_l(x) = 0,&\quad l = 1\dots r.
+    & & g_k(x) \leq 0,&\; k = 1\dots q\,,\\
+    &&& h_l(x) = 0,&\quad l = 1\dots r\,.
   \end{aligned}
@@
 
-Introducing an error threshold $\epsilon$ can convert strict inequalities of the form ${g_k}^{\prime}(x) < 0$ to
-non-strict inequalities of the form $g_k(x) = {g_k}^{\prime}(x) + \epsilon \leq 0$. Adding a tolerance $\tau$ transforms
-equality constraints to a pair of inequalities,
+\textcite{Cop} describe a way to represent non-strict inequality constraints when optimizing using a particle swarm.
+Strict inequalities and equality constraints need to be converted to non-strict inequalities before being represented in
+the problem. Introducing an error threshold $\epsilon$ converts strict inequalities of the form ${g_k}^{\prime}(x) < 0$
+to non-strict inequalities of the form $g_k(x) = {g_k}^{\prime}(x) + \epsilon \leq 0$. A tolerance $\tau$ is used to
+transform equality constraints to a pair of inequalities,
 \begin{equation*}
   \begin{aligned}
-    g_{(q+l)}(x) &= h_l(x) - \tau &\leq 0,&\quad l = 1\dots r,\\
-    g_{(q+r+l)}(x) &= {-}h_l(x) - \tau &\leq 0,&\quad l = 1\dots r.
+    g_{(q+l)}(x) &= h_l(x) - \tau \leq 0,&&\quad l = 1\dots r\,,\\
+    g_{(q+r+l)}(x) &= {-}h_l(x) - \tau \leq 0,&&\quad l = 1\dots r\,.
   \end{aligned}
 \end{equation*}
-Thus, $r$ equality constraints become $2r$ inequality constraints, raising the total number of constraints, denoted by
-$s$, to $s = q + 2r$. For each solution $p_i$, $c_i$ denotes the constraint vector where, $c_{ik} = \max\{g_k(p_i),
-0\},~k=1\dots s$. When $c_{ik} = 0,~\forall k=1\dots s$, the solution $p_i$ lies within the feasible region. When $c_{ik} > 0$,
-the solution $p_i$ violates the $k$\textsuperscript{th} constraint.
+% Cite: PSO Constraints paper.
+Thus, $r$ equality constraints become $2r$ inequality constraints, raising the total number of constraints to $s = q +
+2r$. For each solution $p_i$, $c_i$ denotes the constraint vector where, $c_{ik} = \max\{g_k(p_i), 0\},~k=1\dots s$.
+When $c_{ik} = 0,~\forall k=1\dots s$, the solution $p_i$ lies within the feasible region. When $c_{ik} > 0$, the
+solution $p_i$ violates the $k$\tsup{th} constraint.
 
-Following these guidelines to represent a CO problem, CDHS estimation under CRS becomes,
+Under these guidelines, the following equations represent the CDHS estimation under CRS as a Constrained Optimization problem,
 \begin{equation}\label{eq:cdhscrs}
@@
     & \text{subject to:}
-    &  {-}\phi + \epsilon \leq 0,\quad\phi - 1 + \epsilon \leq 0, &\quad\quad \forall \phi\in\{\alpha,\beta,\gamma,\delta\}\\
-    && (\alpha+\beta-1) - \tau \leq 0,\quad(\gamma+\delta-1) - \tau \leq 0,\\
-    && (1-\alpha-\beta) - \tau \leq 0,\quad(1-\gamma-\delta) - \tau \leq 0,
+    &  {-}\phi + \epsilon \leq 0,\quad\phi - 1 + \epsilon \leq 0\,, &\quad\quad \forall \phi\in\{\alpha,\beta,\gamma,\delta\}\\
+    && (\alpha+\beta-1) - \tau \leq 0,\quad(\gamma+\delta-1) - \tau \leq 0\,,\\
+    && (1-\alpha-\beta) - \tau \leq 0,\quad(1-\gamma-\delta) - \tau \leq 0\,.
   \end{aligned}
 \end{equation}
-but with DRS the last two constraints for $Y_i$ and $Y_s$ are replaced with,
+Under DRS the last two constraints for $Y_i$ and $Y_s$ are replaced with,
 \begin{equation}\label{eq:cdhsdrs}
   \begin{aligned}
-    \alpha + \beta + \epsilon - 1 &\leq 0,\\
-    \gamma + \delta + \epsilon - 1 &\leq 0.
+    \alpha + \beta + \epsilon - 1 &\leq 0\,,\\
+    \gamma + \delta + \epsilon - 1 &\leq 0\,.
   \end{aligned}
 \end{equation}
-
-The CEESA score estimation for DRS is represented as,
+The CEESA score estimation under DRS is represented as,
 \begin{equation}\label{eq:ceesadrs}
@@
     & \text{subject to}
-    &   {-}\phi + \epsilon \leq 0,\quad\phi - 1 + \epsilon \leq 0, &\quad\quad \forall \phi\in\{r,d,t,v,e,\eta\}\\
-    && \rho - 1 \leq 0,\quad \rho - 1 + \epsilon \leq 0,\\
-    && (r+d+t+v+e-1) - \tau \leq 0,\\
-    && (1-r-d-t-v-e) - \tau \leq 0.
+    &   {-}\phi + \epsilon \leq 0,\quad\phi - 1 + \epsilon \leq 0\,, &\quad\quad \forall \phi\in\{r,d,t,v,e,\eta\}\\
+    && \rho - 1 \leq 0,\quad \rho - 1 + \epsilon \leq 0\,,\\
+    && (r+d+t+v+e-1) - \tau \leq 0\,,\\
+    && (1-r-d-t-v-e) - \tau \leq 0\,.
   \end{aligned}
 \end{equation}
-Under CRS there is no need for the parameter $\eta$. The objective function for the problem becomes,
+Under CRS there is no need for the parameter $\eta$ (since $\eta=1$). Thus, the objective function for the problem reduces to,
 \begin{equation}\label{eq:ceesacrs}
   \underset{r,d,t,v,e,\rho}{\text{minimize}}\quad
-  Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\frac{1}{\rho}}.
+  Y = {-}{(r.R^\rho+d.D^\rho+t.{T_s}^\rho+v.{V_e}^\rho+e.E^\rho)}^{\frac{1}{\rho}}\,.
 \end{equation}
@@
 
+% Cite data set.
 The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
@@
 
-Our implemenation used $n=25$ particles to traverse the search space, with learning rates $\lambda_g=0.8$ and
-$\lambda_p=0.2$. It regulated velocity through a friction coefficient of $\mu=0.6$ and upper and lower bounds $\pm1.0$.
-It used an error threshold of $\epsilon=\num{1e-6}$ to convert strict inequalities to non-strict inequalities, and a
-tolerance of $\tau=\num{1e-7}$ to transform an equality constraint to a pair of inequalities. Further implementation
-details are discussed in Appendix~\ref{app:imp}.
+Our implementation used $n=25$ particles to traverse the search space, with learning rates $\lambda_g=0.8$ and
+$\lambda_p=0.2$. It used an inertial weight of $\omega=0.6$ and upper and lower bounds $\pm1.0$. It used an error
+threshold of $\epsilon=\num{1e-6}$ to convert strict inequalities to non-strict inequalities, and a tolerance of
+$\tau=\num{1e-7}$ to transform an equality constraint to a pair of inequalities. Further implementation details are
+discussed in Appendix~\ref{app:imp}.
 
@@
 CEESA scores. All tables also record the number of iterations taken to converge to a stable $\gbest$ value. As seen in
-the tables, although CEESA has 7 parameters and 16 constraints under DRS, PSO takes just over twice the number of
+the tables, although CEESA has 7 parameters and 16 constraints under DRS, PSO takes a little over twice the number of
 iterations to converge as in each step of CDHS, which has 2 parameters and 5 constraints. This is a promising result as
@@
 ($\approx\SI{6}{\minute}\ \SI{11}{\second}$) at \SI{220.36}{\milli\second} per planet, while under CRS it took
-\SI{356.92}{\second} ($\approx\SI{5}{\minute}\ \SI{57}{\second}$) at \SI{212.07}{\milli\second} per planet. This is
-considerably fast considering the number of parameters and constraints for the problem.
+\SI{356.92}{\second} ($\approx\SI{5}{\minute}\ \SI{57}{\second}$) at \SI{212.07}{\milli\second} per planet.
 
@@
 
-\section{Conclusions}
-% - Why is the speed so important?
-% - Parallelizable.
+\section{Conclusion}\label{sec:conc}
 
-\appendix
+\pso\ mainly draws its advantages from being easy to implement and highly parallelizable. The algorithms described in
+Section~\ref{sec:pso} use simple operators and straightforward logic. What is especially noticeable is the lack of the
+need for a gradient, allowing PSO to work in high dimensional search spaces with a large number of constraints.
+Particles of the swarm, in most implementations operate independently during each iteration, their updates can occur
+simultaneously and even asynchronously, yielding much faster execution times than those outlined in
+Section~\ref{sec:results}. However, since strict inequalities and equality constraints are not exactly represented, the
+resulting solution may not be as accurate as direct methods. Despite this, using PSO to calculate the habitability
+scores is beneficial when the number of input parameters are large, which further increases the number of constraints,
+resulting in a model too infeasible for traditional optimization methods.
+
+\printbibliography
+
+\begin{appendices}
+
+  \pagebreak
+  \section{Ensuring Convergence}\label{app:con}
+
+  Although the termination criteria illustrated in Algorithm~\ref{alg:unop}, $|\oldbest-\gbest| < \mathit{threshold}$,
+  is a lucrative solution for unconstrained optimization, it might cause early termination for a constrained problem.
+  This could occur in Algorithm~\ref{alg:cop} as the particles traverse the search space. Due to the way they are
+  updated, all individual particles could move to infeasible regions. There are no updates made to their corresponding
+  $\pbest$s, leading to no change in $\gbest$. Thus, although the swarm has not yet converged to the global optima, it
+  terminates and reports the current $\gbest$ as the optima.
+
+  To prevent this from happening our implementation maintained a counter that was incremented very time $\gbest$ changed
+  by a value less than the threshold it incremented the counter. If the counter reached \num{100}, the algorithm
+  terminated and reported the current value of $\gbest$ as the global optima. If the value of $\gbest$ changed by more
+  than the threshold, it reset the counter to $\num{0}$. Since the function is convex, and all points are initialized to
+  feasible solutions (thus, ensuring that a feasible $\gbest$ exists) there was no need to restrict the total number of
+  possible iterations.
+
+  This explains the spike at \num{0} in Figure~\ref{fig:itercdcrs}. For every exoplanet, the implementation waits a
+  \num{100} iterations for declaring convergence, thus overstating the total iterations by \num{100}. To adjust for this
+  offset, we subtracted \num{100} from the total iterations before constructing the histogram plots in
+  Figures~\ref{fig:itercdcrs},~\ref{fig:itercddrs},~\ref{fig:itercecrs} and~\ref{fig:itercedrs}. Therefore, although the
+  graphs report iterations to convergence for a significant number of planets, only the global best has not shifted
+  locations, but the particles have actually converged to a common point around a region within the threshold of the
+  global best.
 
@@
 
-% Include a description of max_iter and stable_iter.
-Matrices $P$ and $V$ represent current position and velocity, where the $i$\textsuperscript{th} row of each correspond
-to the position and velocity of particle $i$. Each row of the matrix $L$ is the leader for the particle in the
-corresponding row of $P$. The constraint matrix $C$ is constructed by stacking the constraint vectors $c_i$ described in
-Section~\ref{sec:rep}. Let $r',r''$ be two random vectors of length $n$ drawn from the uniform distribution
-${U(0,1)}^n$. Let $X_i$ denote the $i$\textsuperscript{th} row of matrix $X$. The implementation of each iteration while
-updating particles in Algorithm~\ref{alg:cop} reduces to,
+  Most programming languages today have either built-in or external library support for linear algebra routines. The
+  implementation of these libraries are well-tested and efficient, and take advantage of system features such as
+  multiple cores or a general purpose GPU computing device. Taking advantage of these libraries requires representing
+  most entities as either matrices or vectors.
 
-\begin{align*}
-  g' &= f([\gbest])\\
-  L  &= {[\argmin\limits_{pbest_j,\ j = 1\dots n} {\|pbest_j - P_i\|}^2\ \mid\ \forall i = 1\dots n\quad ]}^T\\
-  V  &= \mu.V + \lambda_g\begin{bmatrix}
+  Matrices $P$ and $V$ represent current position and velocity of all particles in the swarm.
+  \begin{alignat}{2}
+    P &= [\ p_{ij}\ &&\mid\ \forall i=1\dots n,\;\forall j=1\dots d\ ]\,,\\
+    V &= [\ v_{ij}\ &&\mid\ \forall i=1\dots n,\;\forall j=1\dots d\ ]\,.
+  \end{alignat}
+  $p_{ij}$ and $v_{ij}$ denote position and velocity of particle $i$ in dimension $d$. Matrices $B$ and $L$ represent
+  the $\pbest$ and $\lbest$ values for the swarm, where,
+  \begin{align}
+    B &= [\ {\pbest}_i\ \mid\ \forall i=1\dots n\ ]\,,\\
+    L &= [\ \argmin_{B_j,\ j = 1\dots n} {\|B_j - P_i\|}^2\ \mid\ \forall i=1\dots n\ ]\,.
+  \end{align}
+  Section~\ref{sec:rep} outlined the use of a constraint vector for describing the feasibility of a solution. This
+  concept generalizes to a matrix $C$, where,
+  \begin{equation}
+    C = [\ c_{ij}\ \mid\ \forall i=1\dots n,\;\forall j=1\dots s\ ]\,.
+  \end{equation}
+  Let $r',r''$ be two random vectors of length $n$ drawn from the uniform distribution ${U(0,1)}^n$. We use the
+  shorthand notation of $X_i$ to denote the $i$\tsup{th} row of some matrix $X$. If $f(X_i)$ is the fitness function and
+  $\mathit{constraints}(X)$ constructs $C$ for solutions $X$, the implementation of each iteration while updating
+  particles in Algorithm~\ref{alg:cop} reduces to,
+  \begin{align}
+    V  =&\ \omega.V + \lambda_g\begin{bmatrix}
       {r_1}'(\gbest - P_1)\\
@@
     \end{bmatrix}\\
-  V  &= [\quad\sgn(v_{ij})*\max\{|v_{max}|,|v_{ij}|\}\ \mid\ \forall i=1\dots n,\forall j=1\dots d\quad]\\
-  P  &= P + V\\
-\end{align*}
+    V  =&\ \left[\ \sgn(v_{ij})*\max\{|v_{max}|,|v_{ij}|\}\ \mid\ \forall i=1\dots n,\;\forall j=1\dots d\ \right]\\
+    P  =&\ P + V\\
+    C  =&\ \mathit{constraints}(P)\\
+    B  =&\ [\ X_i\ \mid\ (\textstyle\sum_{j=1}^s c_{ij} = 0 \to X_i = \argmax\{f(P_i), f(B_i)\})\ \wedge\\
+        &\qquad(\textstyle\sum_{j=1}^s c_{ij} \neq 0 \to X_i = B_i),\ \forall i=1\dots n\ ]\\
+    \gbest =&\ \argmax_{{B}_i} \{\ {f(B_i)}\ \mid\ \forall i=\dots n\ \}
+  \end{align}
+
+\end{appendices}
 

### 2018-01-18 docs/report/report.tex
Fixed an issue with the Bibliography.

--- 2018-01-18 docs/report/draft.tex
@@
-\documentclass[10pt,draft]{article}
+\documentclass[10pt]{article}
 
@@
 \usepackage{booktabs}
+\usepackage{csquotes}
 \usepackage{enumitem}
@@
 \end{equation*}
-% Cite: PSO Constraints paper.
 Thus, $r$ equality constraints become $2r$ inequality constraints, raising the total number of constraints to $s = q +
@@
 
-% Cite data set.
 The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
-Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL). The catalog records observed and
-modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param} describes the
-parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and eccentricity
-are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface temperature
-(\SI{288}{\kelvin}) and eccentricity (\num{0.017}). PHL-EC assumes an Eccentricity of 0 when unavailable. The PHL-EC
-records empty values for planets whose surface temperature is not known. We chose to drop these records from the
-experiment.
+Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL)~\cite{PHL}. The catalog records
+observed and modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param}
+describes the parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and
+eccentricity are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface
+temperature (\SI{288}{\kelvin}) and eccentricity (\num{0.017}). PHL-EC assumes an Eccentricity of 0 when unavailable.
+The PHL-EC records empty values for planets whose surface temperature is not known. We chose to drop these records from
+the experiment.
 
@@
   Matrices $P$ and $V$ represent current position and velocity of all particles in the swarm.
-  \begin{alignat}{2}
+  \begin{alignat*}{2}
     P &= [\ p_{ij}\ &&\mid\ \forall i=1\dots n,\;\forall j=1\dots d\ ]\,,\\
     V &= [\ v_{ij}\ &&\mid\ \forall i=1\dots n,\;\forall j=1\dots d\ ]\,.
-  \end{alignat}
+  \end{alignat*}
   $p_{ij}$ and $v_{ij}$ denote position and velocity of particle $i$ in dimension $d$. Matrices $B$ and $L$ represent
   the $\pbest$ and $\lbest$ values for the swarm, where,
-  \begin{align}
+  \begin{align*}
     B &= [\ {\pbest}_i\ \mid\ \forall i=1\dots n\ ]\,,\\
     L &= [\ \argmin_{B_j,\ j = 1\dots n} {\|B_j - P_i\|}^2\ \mid\ \forall i=1\dots n\ ]\,.
-  \end{align}
+  \end{align*}
   Section~\ref{sec:rep} outlined the use of a constraint vector for describing the feasibility of a solution. This
   concept generalizes to a matrix $C$, where,
-  \begin{equation}
+  \begin{equation*}
     C = [\ c_{ij}\ \mid\ \forall i=1\dots n,\;\forall j=1\dots s\ ]\,.
-  \end{equation}
+  \end{equation*}
   Let $r',r''$ be two random vectors of length $n$ drawn from the uniform distribution ${U(0,1)}^n$. We use the
@@
   particles in Algorithm~\ref{alg:cop} reduces to,
-  \begin{align}
+
+  \begin{align*}
     V  =&\ \omega.V + \lambda_g\begin{bmatrix}
@@
     \gbest =&\ \argmax_{{B}_i} \{\ {f(B_i)}\ \mid\ \forall i=\dots n\ \}
-  \end{align}
+  \end{align*}
 

### 2018-01-19 docs/report/report.tex
Finished the report. Mailed.

@@
 planets~\cite{Kepler,Presto} has been an international venture since Frank Drake's attempt with Project Ozma in the
-mid-20th century~\cite{Ozma}, demanding large investments in cost and effort. \textcite{Exo1} confirmed the first
-exoplanet in 1991. This marked the start of a trend that has lasted 25 years and yielded over 3,700 confirmed
-exoplanets. There have been attempts to assess the habitability of these planets and to assign a score based on their
-similarity to Earth. Two such habitability scores are the Cobb-Douglas Habitability Score (CDHS)~\cite{Bora,Saha} and
-the Constant Elasticity Earth Similarity Approach (CEESA) score. Estimating these scores involves maximizing a
-production function while observing a set of constraints on the input variables.
+mid-20th century~\cite{Ozma}. \textcite{Exo1} confirmed the first exoplanet in 1991. This marked the start of a trend
+that has lasted 25 years and yielded over 3,700 confirmed exoplanets. There have been attempts to assess the
+habitability of these planets and to assign a score based on their similarity to Earth. Two such habitability scores are
+the Cobb-Douglas Habitability Score (CDHS)~\cite{Bora,Saha} and the Constant Elasticity Earth Similarity Approach
+(CEESA) score. Estimating these scores involves maximizing a production function while observing a set of constraints on
+the input variables.
 
@@
 for non-polynomial functions in high-dimensional search spaces. Further, subjecting the input variables to constraints,
-as needed by CDHS and CEESA, are not always straightforward to represent within the model. This paper presents an
+as needed by CDHS and CEESA, are not always straightforward to represent within the model. This paper details an
 approach to Constrained Optimization (CO) using the swarm intelligence metaheuristic. \pso\ (PSO) is a method for
-optimizing a continuous function that does away with the need for a gradient. It employs a large number of particles
-that traverse the search space converging toward a global best solution encountered by at least one of the
-particles~\cite{PSO1,PSO2}.
+optimizing a continuous function that does away with the need for calculating the gradient. It employs a large number of
+randomly initialized particles that traverse the search space, eventually converging at a global best solution
+encountered by at least one particle~\cite{PSO1,PSO2}.
 
@@
 lucrative solution where computational resources are at a premium. Its implementation is highly parallelizable. It
-scales with the dimensionality of the search space. The standard PSO algorithm does not deal with constraints but
-through variations in initializing and updating particles constraints are straightforward to represent and adhere to, as
-seen in Section~\ref{subsec:copso}. \textcite{Poli1,Poli2} carried out extensive surveys on the applications of PSO
+scales with the dimensionality of the search space. The standard PSO algorithm does not deal with constraints but,
+through variations in initializing and updating particles, constraints are straightforward to represent and adhere to,
+as seen in Section~\ref{subsec:copso}. \textcite{Poli1,Poli2} carried out extensive surveys on the applications of PSO,
 reporting uses in Communication Networks, Machine Learning, Design, Combinatorial Optimization and Modeling, among
@@
 
-This paper demonstrates the applicability of \pso\ in estimating CDHS and CEESA scores of an exoplanet by maximizing
+This paper demonstrates the applicability of \pso\ in estimating the CDHS and CEESA scores of an exoplanet by maximizing
 their respective production functions (discussed in Sections~\ref{subsec:cdhs} and~\ref{subsec:ceesa}). CDHS considers
@@
 where, $R$, $D$, $V_e$ and $T_s$ are density, radius, escape velocity and surface temperature respectively. $\alpha$,
-$\beta$, $\gamma$ and $\delta$ are the elasticity coefficients subject to,
-\begin{equation}
-  0 < \alpha,\beta,\gamma,\delta < 1\,.
-\end{equation}
-Equations~\ref{eq:cdhsi} and~\ref{eq:cdhss} are convex under either Constant Returns to Scale (CRS) or Decreasing
-Returns to Scale (DRS) marked by two constraints on the elasticity coefficients,
-\begin{subequations}
-  \begin{alignat}{3}
-    \text{CRS:} & \quad\alpha+\beta = 1,&\quad\gamma+\delta = 1\,,\\
-    \text{DRS:} & \quad\alpha+\beta < 1,&\quad\gamma+\delta < 1\,.
-  \end{alignat}
-\end{subequations}
-The final CDHS is the convex combination of the interior and surface CDHS values as given by,
+$\beta$, $\gamma$ and $\delta$ are the elasticity coefficients subject to $0 < \alpha,\beta,\gamma,\delta < 1$.
+Equations~\ref{eq:cdhsi} and~\ref{eq:cdhss} are convex under either Constant Returns to Scale (CRS), when
+$\alpha+\beta=1$ and $\gamma+\delta=1$, or Decreasing Returns to Scale (DRS), when $\alpha+\beta<1$ and
+$\gamma+\delta<1$. The final CDHS is the convex combination of the interior and surface scores as given by,
 \begin{equation}
@@
 \end{equation}
-where, $E$ is the fifth parameter denoting Orbital Eccentricity. The value of $\rho$ lies within $0<\rho\leq 1$.
-The coefficients ($r$, $d$, $t$, $v$ and $e$) are constrained by,
-\begin{subequations}
-  \begin{align}
-      0 < r,d,t,v,e < 1\,,\\
-      r+d+t+v+e = 1\,.
-  \end{align}
-\end{subequations}
-The value of $\eta$ is constrained by the scale of production used,
-\begin{subequations}
-  \begin{alignat}{3}
-    \text{CRS:} & \quad 0 < \eta \leq 1\,,\\
-    \text{DRS:} & \quad \eta = 1\,.
-  \end{alignat}
-\end{subequations}
+where, $E$ is the fifth parameter denoting Orbital Eccentricity. The value of $\rho$ lies within $0<\rho\leq 1$. The
+coefficients $r$, $d$, $t$, $v$ and $e$ lie in $(0,1)$ and sum to \num{1}, $r+d+t+v+e = 1$. The value of $\eta$ is
+constrained by the scale of production used, $0 < \eta < 1$ under DRS and $\eta=1$ under CRS.
 
@@
 
-% Cite: PSO paper.
 Particle Swarm Optimization (PSO)~\cite{PSO1} is a biologically inspired metaheuristic for finding the global minima of
-a function.  Traditionally designed for unconstrained inputs, it works by iteratively converging a population of
-randomly initialized solutions, called particles, toward a globally optimal solution. Each particle in the population
-keeps track of its current position and the best solution it has encountered so far, called $\pbest$. Each particle also
-has an associated randomized velocity used to traverse the search space. The swarm keeps track of the overall best
-solution, called $\gbest$. Each iteration of the swarm updates the velocity of the particle towards its $\pbest$ and the
-$\gbest$ values.
+a function. Traditionally designed for unconstrained inputs, it works by iteratively converging a population of randomly
+initialized solutions, called particles, toward a globally optimal solution. Each particle in the population keeps track
+of its current position and the best solution it has encountered, called $\pbest$. Each particle also has an associated
+velocity used to traverse the search space. The swarm keeps track of the overall best solution, called $\gbest$. Each
+iteration of the swarm updates the velocity of the particle towards its $\pbest$ and the $\gbest$ values.
 
@@
 $\pbest$ to the new position. Once the algorithm has updated all particles, it updates $\gbest$ to the new overall best
-position. A suitable termination criteria for the swarm, under convex optimization, is  when the $\gbest$ position has
-not changed by the end of the iteration.
+position. A suitable termination criteria for the swarm, under convex optimization, is to terminate when $\gbest$ has
+not changed by the end of an iteration.
 
@@
 
-Under these guidelines, the following equations represent the CDHS estimation under CRS as a Constrained Optimization problem,
+Under these guidelines, the representation of CDHS estimation (described in Section~\ref{subsec:cdhs}) under CRS as a
+CO problem is given by,
+
 \begin{equation}\label{eq:cdhscrs}
@@
 \end{equation}
-The CEESA score estimation under DRS is represented as,
+
+The representation of CEESA score estimation (described in Section~\ref{subsec:ceesa}) under DRS as a CO problem is
+given by,
 \begin{equation}\label{eq:ceesadrs}
@@
 
-The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
-Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL)~\cite{PHL}. The catalog records
-observed and modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param}
-describes the parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and
-eccentricity are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface
-temperature (\SI{288}{\kelvin}) and eccentricity (\num{0.017}). PHL-EC assumes an Eccentricity of 0 when unavailable.
-The PHL-EC records empty values for planets whose surface temperature is not known. We chose to drop these records from
-the experiment.
+The data set used for estimating the Habitability Scores of exoplanets was the Confirmed Exoplanets Catalog maintained
+by the Planetary Habitability Laboratory (PHL)~\cite{PHL}. The catalog records observed and modeled parameters for
+exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param} describes the parameters from the PHL
+Exoplanets' Catalog (PHL-EC) used for the experiment. Since surface temperature and eccentricity are not recorded in
+Earth Units, we normalized these values by dividing them with Earth's surface temperature (\SI{288}{\kelvin}) and
+eccentricity (\num{0.017}). PHL-EC assumes an Eccentricity of 0 when unavailable. The PHL-EC records empty values for
+planets whose surface temperature is not known. We chose to drop these records from the experiment.
 
 Our implementation used $n=25$ particles to traverse the search space, with learning rates $\lambda_g=0.8$ and
-$\lambda_p=0.2$. It used an inertial weight of $\omega=0.6$ and upper and lower bounds $\pm1.0$. It used an error
+$\lambda_p=0.2$. It used an inertial weight of $\omega=0.6$ and upper and lower bounds $\pm1.0$. We used an error
 threshold of $\epsilon=\num{1e-6}$ to convert strict inequalities to non-strict inequalities, and a tolerance of
 $\tau=\num{1e-7}$ to transform an equality constraint to a pair of inequalities. Further implementation details are
-discussed in Appendix~\ref{app:imp}.
+discussed in the Appendix.
 
@@
 estimate the CDHS under CRS for \num{1683} exoplanets, at an average of \SI{198.11}{\milli\second} for each planet for
-each part of the CDHS.\@ For CDHS under DRS, it took \SI{638.69}{\second} ($\approx\SI{10}{\minute}\ \SI{39}{\second}$)
-at an average of \SI{189.75}{\milli\second} for each part of the CDHS.\@ The CEESA estimates, requiring one calculation,
-took a little over half the total time to execute. Under DRS it took a total of \SI{370.86}{\second}
-($\approx\SI{6}{\minute}\ \SI{11}{\second}$) at \SI{220.36}{\milli\second} per planet, while under CRS it took
-\SI{356.92}{\second} ($\approx\SI{5}{\minute}\ \SI{57}{\second}$) at \SI{212.07}{\milli\second} per planet.
+each individual score (interior and surface) of the CDHS.\@ For CDHS under DRS, it took \SI{638.69}{\second}
+($\approx\SI{10}{\minute}\ \SI{39}{\second}$) at an average of \SI{189.75}{\milli\second} for each part of the CDHS.\@
+The CEESA calculations, requiring a single estimate, took a little over half the CDHS execution time to run. Under DRS
+it took a total of \SI{370.86}{\second} ($\approx\SI{6}{\minute}\ \SI{11}{\second}$) at \SI{220.36}{\milli\second} per
+planet, while under CRS it took \SI{356.92}{\second} ($\approx\SI{5}{\minute}\ \SI{57}{\second}$) at
+\SI{212.07}{\milli\second} per planet.
 
@@
 Section~\ref{sec:pso} use simple operators and straightforward logic. What is especially noticeable is the lack of the
-need for a gradient, allowing PSO to work in high dimensional search spaces with a large number of constraints.
-Particles of the swarm, in most implementations operate independently during each iteration, their updates can occur
-simultaneously and even asynchronously, yielding much faster execution times than those outlined in
-Section~\ref{sec:results}. However, since strict inequalities and equality constraints are not exactly represented, the
-resulting solution may not be as accurate as direct methods. Despite this, using PSO to calculate the habitability
-scores is beneficial when the number of input parameters are large, which further increases the number of constraints,
-resulting in a model too infeasible for traditional optimization methods.
+need for a gradient, allowing PSO to work in high dimensional search spaces with a large number of constraints,
+precisely what is needed in a potential Habitability score estimate. Further, particles of the swarm, in most
+implementations operate independently during each iteration, their updates can occur simultaneously and even
+asynchronously, yielding much faster execution times than those outlined in Section~\ref{sec:results}. However, since
+strict inequalities and equality constraints are not exactly represented, the resulting solution may not be as accurate
+as direct methods. Despite this, using PSO to calculate the habitability scores is beneficial when the number of input
+parameters are large, which further increases the number of constraints, resulting in a model too infeasible for
+traditional optimization methods.
 
@@
 
-  \pagebreak
   \section{Ensuring Convergence}\label{app:con}
@@
 
-  To prevent this from happening our implementation maintained a counter that was incremented very time $\gbest$ changed
-  by a value less than the threshold it incremented the counter. If the counter reached \num{100}, the algorithm
-  terminated and reported the current value of $\gbest$ as the global optima. If the value of $\gbest$ changed by more
-  than the threshold, it reset the counter to $\num{0}$. Since the function is convex, and all points are initialized to
-  feasible solutions (thus, ensuring that a feasible $\gbest$ exists) there was no need to restrict the total number of
-  possible iterations.
+  To prevent this from happening our implementation maintained a counter that it incremented every time $\gbest$ changed
+  by a value less than the threshold. When the counter reached \num{100}, the algorithm terminated and reported the
+  current value of $\gbest$ as the global optima. If the value of $\gbest$ changed by more than the threshold, it reset
+  the counter to $\num{0}$. Since the function is convex, and all points are initialized to feasible solutions (thus,
+  ensuring that a feasible $\gbest$ has been encountered) there was no need to restrict the total number of possible
+  iterations.
 
@@
   Figures~\ref{fig:itercdcrs},~\ref{fig:itercddrs},~\ref{fig:itercecrs} and~\ref{fig:itercedrs}. Therefore, although the
-  graphs report iterations to convergence for a significant number of planets, only the global best has not shifted
-  locations, but the particles have actually converged to a common point around a region within the threshold of the
-  global best.
+  graphs report iterations to convergence for a significant number of planets, only $\gbest$ has not shifted, the
+  particles have actually converged to a common point around a region within the threshold of the global best.
 
@@
   shorthand notation of $X_i$ to denote the $i$\tsup{th} row of some matrix $X$. If $f(X_i)$ is the fitness function and
-  $\mathit{constraints}(X)$ constructs $C$ for solutions $X$, the implementation of each iteration while updating
-  particles in Algorithm~\ref{alg:cop} reduces to,
-
+  $\mathit{constraints}(X)$ constructs $C$ for solutions $X$, the update of each particle in Algorithm~\ref{alg:cop} is
+  implemented as,
   \begin{align*}
@@
         &\qquad(\textstyle\sum_{j=1}^s c_{ij} \neq 0 \to X_i = B_i),\ \forall i=1\dots n\ ]\\
-    \gbest =&\ \argmax_{{B}_i} \{\ {f(B_i)}\ \mid\ \forall i=\dots n\ \}
   \end{align*}
+  Finally, after all particles are updated, $\gbest$ is updated according to,
+  \begin{equation*}
+    \gbest = \argmax_{{B}_i} \{\ {f(B_i)}\ \mid\ \forall i=\dots n\ \}\,.
+  \end{equation*}
 
